---
title: "Homework1_BST226"
author: "Takuto Yoshida"
date: "2023-01-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question1
Calculate the mean salary for male and female faculty. What does the comparison of the means suggest about the relationship between sex and faculty salaries? 
```{r}
# import the data
df <- read.table("/Users/yoshidatakuto/Dropbox/HSPH/MPH-CLE/Spring/BST226/Homework/salary.txt")
summary(df)

# Assigning column names to a data frame
colnames(df) <- c("ID", "sex", "years", "rank", "salary")

# Calculating the average salary by sex
library(dplyr)
df %>% 
  group_by(sex) %>%
  summarize(mean_salary = mean(salary))
```

# Question2
Using salary as the outcome variable, ﬁt a simple linear regression model with sex as the only predictor. Interpret the estimated intercept and coefﬁcient for sex.                               
Note: sex is coded 0=Male, 1=Female. 
```{r}
# Fitting a simple linear regression model
lm_mod_1 <- lm(salary ~ sex, data = df)
summary(lm_mod_1)
```
# Question5
Next, consider the effects of faculty rank and years in rank. Recall that rank is categorical with 3 levels; use the ﬁrst level (1=Assistant Professor) as the “reference” and construct dummy or indicator variables for rank . Using salary as the response variable, ﬁt a multiple linear regression model with sex, rank, and years in rank as predictors and present the table of regression coefficients, SEs, t-statistics, and p-values. 
```{r}
# Creating and set 2 new columns to zero
df$rank2 = 0
df$rank3 = 0

# Where rank=2, replace the new rank2 column with value 1
df$rank2[which(df$rank == 2)] = 1

# Where rank=3, replace the new rank3 column with value 1
df$rank3[which(df$rank == 3)] = 1

# Fitting the multivariable linear moder with dummy variable
mod_q5 <- lm(salary ~ rank2 + rank3 + sex + years, data = df)
summary(mod_q5)

```
# Question8
```{r}
# Method1(Simple Diaganostic): Plot salary and years in rank
plot(x = df$years, y = df$salary, xlab="years in rank", ylab="salary", main="Question8 Answe")

# Method2(Simple Diagnostic): Using product term of years
mod_8a <- lm(salary ~ years + I(years*years) + I(years*years*years), data = df)
summary(mod_8a)

mod_8b <- lm(salary ~ years + I(years*years) , data = df)
summary(mod_8b)

mod_8c <- lm(salary ~ years, data = df)
summary(mod_8c)

# Method3 (Intermediate Diagnostic); Plot model residuals and years.
plot(x=residuals(mod_q5), y=df$years, xlab="RSE", ylab="years", main="Scatter plot of the residuals versus years in rank")

# Method4 
# Creating the residuals
hist(residuals(mod_q5))

# Creating the QQ plot of mod_q5 residuals
library(car)
qqPlot(mod_q5)

# Doing Shapiro-Wilk test
shapiro.test(df$salary)
shapiro.test(residuals(mod_q5))
```
# Question 9
The Dean for Faculty Affairs conjectures that the effect of years in rank  on salaries may be different for male and female faculty. Construct a statistical test of this hypothesis. What do you conclude? 
```{r}
# Create the interaction term
df$sexyrs <- df$sex * df$years

# Fitting the new model by using the interaction term
mod_q9 <- lm(salary ~ sex + years + sexyrs, data = df)
summary(mod_q9)
```















