---
title: "covid_surge"
author: "Takuto Yoshida"
date: "2023-02-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Install package

```{r}
if (!require(readxl)) {install.packages("readxl")}
if (!require(ggplot2)) {install.packages("ggplot2")}
if (!require(dplyr)) {install.packages("dplyr")}
if (!require(car)) {install.packages("car")}
if (!require(car)) {install.packages("tidyr")}
if (!require(car)) {install.packages("stringi")}
if (!require(lubridate)) {install.packages("lubridate")}
if (!require(mice)) {install.packages("mice")}
if (!require(table1)) {install.packages("table1")}
if (!require(geepack)) {install.packages("geepack")}
library(geepack)
library(lubridate)
library(table1)
library(ggthemes)
library(ggrepel)
library(dplyr)
```

# Create the dataset

```{r}
covsurge <- read.csv("/Users/yoshidatakuto/Dropbox/HSPH/MPH-CLE/EPI 946/Dataset/nsqip_op.csv")
vaccine <- read.csv("/Users/yoshidatakuto/Dropbox/HSPH/MPH-CLE/EPI 946/Dataset/Final version - COVID vaccination prior to surgery 2019 to 2022 (1).csv")

# Remove duplicates from the "vaccine" dataset
vaccine <- vaccine %>%
  distinct(PATIENT_DISPLAY_ID, .keep_all = TRUE)

covsurge <- covsurge %>%
  left_join(vaccine, by = c("IDN"="PATIENT_DISPLAY_ID"))
```

## Operation Day

```{r}
# Separate Operation.Date into Year, Month, and Day
class(covsurge$Operation.Date)
covsurge <- covsurge %>%
  mutate(opedate = as.Date(Operation.Date, format ="%m/%d/%Y"))
class(covsurge$opedate)

#Extract data after March 1, 2020.
covsurge <- covsurge %>%
   filter(opedate >= as.Date("2020-03-01"))
```

# COVID Surge

```{r}
# Define a date range during the surge period
## December11, 2020 - March1, 2021
## December27, 2021 - January31, 2022
surge_start_dates <- c("2020-12-11", "2021-12-27")
surge_end_dates <- c("2021-03-01", "2022-01-31")

# Assign 1 to the data corresponding to the surge period 
covsurge$surge <- 0
for (i in 1:length(surge_start_dates)) {
  covsurge$surge[covsurge$opedate >= as.Date(surge_start_dates[i]) & 
                  covsurge$opedate <= as.Date(surge_end_dates[i])] <- 1
}

# Check for correct assignment
# subset(covsurge, surge == 0 ) $opedate
# subset(covsurge, surge == 1 ) $opedate
```

## Age

```{r}
# countinuous
covsurge$age <- covsurge$Age.at.Time.of.Surgery

# Age category
covsurge$age_cat <- ifelse(covsurge$age < 40, 1, 
                           ifelse(covsurge$age <=50, 2,
                                  ifelse(covsurge$age <= 60, 3, 
                                         ifelse(covsurge$age <= 70, 4, 5))))
summary(covsurge$age)
table(covsurge$age_cat)
sum(is.na(covsurge$age)) # No missing value
covsurge$age_cat <- as.factor(covsurge$age_cat)
```

## Sex

```{r}
covsurge$sex <- ifelse(covsurge$Sex == "Male", 1, 0) # Female:0, Male:1
covsurge$sex <- as.factor(covsurge$sex)
table(covsurge$sex)
sum(is.na(covsurge$sex)) # No missing value
covsurge$sex <- as.factor(covsurge$sex)
```

## Race

```{r}
table(covsurge$Race)
covsurge$Race[covsurge$Race ==""] <- NA

# Race=0: Black or African American, Race=1: White, Race=2: Others
covsurge$race <- ifelse(covsurge$Race=="Black or African American", 0,
                        ifelse(covsurge$Race=="White", 1,2))

covsurge$race <- as.factor(covsurge$race)
sum(is.na(covsurge$race))
table(covsurge$race, useNA="always")  # 24 missing values
```

## BMI

```{r}
covsurge$bmi <- ifelse(covsurge$BMI < 18.5, 0,  # BMI<18.5: 0
                           ifelse(covsurge$BMI <25, 1, # BMI 18.5-25: 1
                                  ifelse(covsurge$BMI < 30, 2, # BMI 25-30: 2
                                         ifelse(covsurge$BMI >= 30, 3, NA)))) # BMI 30-: 3, missing value: NA 
table(covsurge$bmi, useNA="always")
sum(is.na(covsurge$bmi))  
covsurge$bmi <- as.factor(covsurge$bmi) # 47 missing value
```

## Functional Status

```{r}
table(covsurge$Functional.Heath.Status)
covsurge$Functional.Heath.Status[covsurge$Functional.Heath.Status ==""] <- NA

# func=0: Independent, func=1: Partially Dependent, func=2: Tottally Dependent, func=3: Unknown
covsurge$func <- ifelse(covsurge$Functional.Heath.Status=="Independent", 0,
                        ifelse(covsurge$Functional.Heath.Status=="Partially Dependent", 1, 
                        ifelse(covsurge$Functional.Heath.Status=="Totally Dependent", 2,
                        ifelse(covsurge$Functional.Heath.Status=="Unknown",3,NA))))

sum(is.na(covsurge$Functional.Heath.Status))
table(covsurge$func, useNA="always") 
covsurge$func <- as.factor(covsurge$func) # two missing value
```

## Socioeconomic Status

```{r}
table(covsurge$Primary.Payor.Status)

# ses=0: medicaid, ses=1: medicare, ses=2: private insurance, ses=3: self-pay, ses=4: others

covsurge$ses <- ifelse(covsurge$Primary.Payor.Status=="Medicaid", 0,
                        ifelse(covsurge$Primary.Payor.Status=="Medicare", 1, 
                        ifelse(covsurge$Primary.Payor.Status=="Private Insurance", 2,
                        ifelse(covsurge$Primary.Payor.Status=="Self-Pay",3,4))))

sum(is.na(covsurge$ses))
table(covsurge$ses, useNA="always") 
covsurge$ses <- as.factor(covsurge$ses) # two missing value
```

## Emergency

```{r}
table(covsurge$Emergency.Case)

# emergency=0: Elective surgery, emergency=1: Emergency surgery
covsurge$emergency <- ifelse(covsurge$Emergency.Case=="No", 0,
                             ifelse(covsurge$Emergency.Case=="Yes", 1, NA))
sum(is.na(covsurge$emergency))
table(covsurge$emergency, useNA="always")
covsurge$emergency <- as.factor(covsurge$emergency) # 5612 missing value
```

More than 50% are missing values, so they are not incorporated into the final model. This will be added to Limitation.

## ASA-PS

```{r}
table(covsurge$ASA.Classification)

# ASA-PS category
covsurge$asa <- ifelse(covsurge$ASA.Classification=="ASA 1 - No Disturb", 1,
                             ifelse(covsurge$ASA.Classification=="ASA 2 - Mild Disturb", 2,
                                    ifelse(covsurge$ASA.Classification=="ASA 3 - Severe Disturb", 3,
                                           ifelse(covsurge$ASA.Classification=="ASA 4 - Life Threat", 4,
                                                  ifelse(covsurge$ASA.Classification=="ASA 5 - Moribund", 5, NA)))))
sum(is.na(covsurge$asa))
table(covsurge$asa, useNA="always") 
covsurge$asa <- as.factor(covsurge$asa) # 37 missing values
```

## DM

```{r}
table(covsurge$Diabetes.Mellitus)

# DM category: 0=No, 1=Non-Insulin, 2=Insulin
covsurge$dm <- ifelse(covsurge$Diabetes.Mellitus=="No", 0,
                             ifelse(covsurge$Diabetes.Mellitus=="Non-Insulin", 1,
                                    ifelse(covsurge$Diabetes.Mellitus=="Insulin", 2, NA)))
sum(is.na(covsurge$dm))
table(covsurge$dm, useNA="always") 
covsurge$dm <- as.factor(covsurge$dm) # 2 missing values
```

## HTN

```{r}
table(covsurge$Hypertension.requiring.medication)

# HTN category: 0=No, 1=Yes
covsurge$htn <- ifelse(covsurge$Hypertension.requiring.medication=="No", 0,
                             ifelse(covsurge$Hypertension.requiring.medication=="Yes", 1, NA))
sum(is.na(covsurge$htn))
table(covsurge$htn, useNA="always") 
covsurge$htn <- as.factor(covsurge$htn) # 2 missing values
covsurge$htn <- as.factor(covsurge$htn)
```

## Immunosuppressive Therapy

```{r}
table(covsurge$Immunosuppressive.Therapy)

# Immunosuppressive category: 0=No, 1=Yes
covsurge$imsp <- ifelse(covsurge$Immunosuppressive.Therapy=="No", 0,
                             ifelse(covsurge$Immunosuppressive.Therapy=="Yes", 1, NA))
sum(is.na(covsurge$imsp))
table(covsurge$imsp, useNA="always") 
covsurge$imsp <- as.factor(covsurge$imsp) # 2 missing values
```

## Ascites

```{r}
table(covsurge$Ascites)

# Ascites category: 0=No, 1=Yes
covsurge$ascites <- ifelse(covsurge$Ascites=="No", 0,
                             ifelse(covsurge$Ascites=="Yes", 1, NA))
sum(is.na(covsurge$ascites))
table(covsurge$ascites, useNA="always") 
covsurge$ascites <- as.factor(covsurge$ascites) # 2 missing values
```

## Systemic sepsis within 48 h before surgery

```{r}
table(covsurge$Sepsis..SIRS.Sepsis.Septic.shock...48h.)

# Sepsis category: 0=None, 1=Sepsis, 2=Septic shock, 3=SIRS
covsurge$sepsis <- ifelse(covsurge$Sepsis..SIRS.Sepsis.Septic.shock...48h.=="None", 0,
                             ifelse(covsurge$Sepsis..SIRS.Sepsis.Septic.shock...48h.=="Sepsis", 1,
                                    ifelse(covsurge$Sepsis..SIRS.Sepsis.Septic.shock...48h.=="Septic shock", 3,
                                           ifelse(covsurge$Sepsis..SIRS.Sepsis.Septic.shock...48h.=="SIRS", 4, NA))))
sum(is.na(covsurge$sepsis))
table(covsurge$sepsis, useNA="always") 
covsurge$sepsis <- as.factor(covsurge$sepsis) # 2 missing values
```

## Ventilator dependency

```{r}
table(covsurge$Ventilator.Dependent)

# Ventilator dependency category: 0=No, 1=Yes
covsurge$venti <- ifelse(covsurge$Ventilator.Dependent=="No", 0,
                             ifelse(covsurge$Ventilator.Dependent=="Yes", 1, NA))
sum(is.na(covsurge$venti))
table(covsurge$venti, useNA="always") 
covsurge$venti <- as.factor(covsurge$venti) # 2 missing values
```

## Disseminated cancer

```{r}
table(covsurge$Disseminated.Cancer)

# Disseminated.Cancer category: 0=No, 1=Yes
covsurge$cancer <- ifelse(covsurge$Disseminated.Cancer=="No", 0,
                             ifelse(covsurge$Disseminated.Cancer=="Yes", 1, NA))
sum(is.na(covsurge$cancer))
table(covsurge$cancer, useNA="always") 
covsurge$cancer <- as.factor(covsurge$cancer) #2 missing values
```

## Congestive heart failure

```{r}
table(covsurge$Heart.Failure)

# Heart failure category: 0=No, 1=Yes
covsurge$chf <- ifelse(covsurge$Heart.Failure=="No", 0,
                             ifelse(covsurge$Heart.Failure=="Yes", 1, NA))
sum(is.na(covsurge$chf))
table(covsurge$chf, useNA="always") #2 missing values
covsurge$chf <- as.factor(covsurge$chf)
```

## Dyspnea

```{r}
table(covsurge$Dyspnea)

# Dyspnea category: 0=No, 1=Moderate Exertion, 2=AtRest
covsurge$dyspnea <- ifelse(covsurge$Dyspnea=="No", 0,
                             ifelse(covsurge$Dyspnea=="Moderate Exertion", 1,
                                    ifelse(covsurge$Dyspnea=="AtRest", 2, NA)))
covsurge$dyspnea[covsurge$dyspnea =="NA"] <- NA
sum(is.na(covsurge$dyspnea))
table(covsurge$dyspnea, useNA="always") 
covsurge$dyspnea <- as.factor(covsurge$dyspnea) #5620 missing values
```

This is also missing in about half of the cases. However, since this one is not incorporated into the final model, no assignment of missing values is made.

## Current smoker within 1y

```{r}
table(covsurge$Current.Smoker.within.1.year)

# smoking category: 0=No, 1=Yes
covsurge$smk <- ifelse(covsurge$Current.Smoker.within.1.year=="No", 0,
                             ifelse(covsurge$Current.Smoker.within.1.year=="Yes", 1, NA))
sum(is.na(covsurge$smk))
table(covsurge$smk, useNA="always") #2 missing values
covsurge$smk <- as.factor(covsurge$smk)
```

## COPD

```{r}
table(covsurge$History.of.Severe.COPD)

# COPD category: 0=No, 1=Yes
covsurge$copd <- ifelse(covsurge$History.of.Severe.COPD=="No", 0,
                             ifelse(covsurge$History.of.Severe.COPD=="Yes", 1, NA))
sum(is.na(covsurge$copd))
table(covsurge$copd, useNA="always") #2 missing values
covsurge$copd <- as.factor(covsurge$copd)
```

## Dialysis

```{r}
table(covsurge$Preop.Dialysis)

# Dialysis category: 0=No, 1=Yes
covsurge$dialysis <- ifelse(covsurge$Preop.Dialysis=="No", 0,
                             ifelse(covsurge$Preop.Dialysis=="Yes", 1, NA))
sum(is.na(covsurge$dialysis))
table(covsurge$dialysis, useNA="always") #2 missing values
covsurge$dialysis <- as.factor(covsurge$dialysis)
```

## Acute renal failure

```{r}
table(covsurge$Preop.Acute.Kidney.Injury)

# AKI category: 0=No, 1=Yes
covsurge$aki <- ifelse(covsurge$Preop.Acute.Kidney.Injury=="No", 0,
                             ifelse(covsurge$Preop.Acute.Kidney.Injury=="Yes", 1, NA))
sum(is.na(covsurge$aki))
table(covsurge$aki, useNA="always") #2 missing values
covsurge$aki <- as.factor(covsurge$aki)
```

## COVID-19 infection at the time of surgery

```{r}
table(covsurge$Preop.COVID.19.Diagnosis)

# Preop covid category: 0=No, 1=Yes, lab-confirmed diagnosis (or ICD-10 code U07.1), 2=Yes, suspected diagnosis (or ICD-10 code U07.2)
covsurge$covid <- ifelse(covsurge$Preop.COVID.19.Diagnosis=="No", 0,
                             ifelse(covsurge$Preop.COVID.19.Diagnosis=="Yes, lab-confirmed diagnosis (or ICD-10 code U07.1)", 1,
                                    ifelse(covsurge$Preop.COVID.19.Diagnosis=="Yes, suspected diagnosis (or ICD-10 code U07.2)", 2, NA)))

covsurge$covid[covsurge$covid ==""] <- NA
sum(is.na(covsurge$covid))
table(covsurge$covid, useNA="always") 
covsurge$covid <- as.factor(covsurge$covid) # 3 missing values
```

## COVID-19 Vaccination

```{r}
# The dataset has only 2626 patients. Maybe it will related to strong bias and remove it.
table(covsurge$Vaccinated.prior.to.Surgery.)

# Filter rows with missing values in Vaccinated.prior.toSurgery
missing_vaccinated_data <- covsurge %>%
  filter(is.na(Vaccinated.prior.to.Surgery.)) %>%
  select(opedate)

# Display the missing_vaccinated_data
print(missing_vaccinated_data) # All missing values were prior to Sep. 15, 2020.

# pat_vac category: 0=No, 1=Yes
covsurge <- covsurge %>%
  mutate(pat_vac = ifelse(Vaccinated.prior.to.Surgery.=="Yes", 1,
                          ifelse(Vaccinated.prior.to.Surgery.=="No"|is.na(Vaccinated.prior.to.Surgery.),0, NA)))
covsurge$pat_vac[is.na(covsurge$pat_vac)] <-0
sum(is.na(covsurge$pat_vac))
table(covsurge$pat_vac, useNA="always") #2 missing values
covsurge$pat_vac <- as.factor(covsurge$pat_vac)
```

# Operation
## Type of surgery

```{r}
table(covsurge$Surgical.Specialty)

# Type of surgery category: 0=General Surgery, 1=Neurosurgery, 2=Otolaryngology (ENT), 3=Plastics, 4=Thoracic, 5=Urology, 6=Vascular
covsurge$surg_type <- ifelse(covsurge$Surgical.Specialty=="General Surgery", 0,
                             ifelse(covsurge$Surgical.Specialty=="Neurosurgery", 1,
                                    ifelse(covsurge$Surgical.Specialty=="Otolaryngology (ENT)", 2,
                                           ifelse(covsurge$Surgical.Specialty=="Plastics", 3,
                                                 ifelse(covsurge$Surgical.Specialty=="Thoracic", 4,
                                                        ifelse(covsurge$Surgical.Specialty=="Urology", 5,
                                                               ifelse(covsurge$Surgical.Specialty=="Vascular", 6,NA)))))))

sum(is.na(covsurge$surg_type))
table(covsurge$surg_type, useNA="always") # no missing values
covsurge$surg_type <- as.factor(covsurge$surg_type)
```

## Type of Anesthesia

```{r}
table(covsurge$Principal.Anesthesia.Technique)

# Type of surgery category: 0=No, 1=Yes, lab-confirmed diagnosis (or ICD-10 code U07.1), 2=Yes, suspected diagnosis (or ICD-10 code U07.2)
covsurge$anes <- ifelse(covsurge$Principal.Anesthesia.Technique=="Regional", 0,
                             ifelse(covsurge$Principal.Anesthesia.Technique=="General", 1,
                                    ifelse(covsurge$Principal.Anesthesia.Technique=="Epidural", 2,
                                           ifelse(covsurge$Principal.Anesthesia.Technique=="Spinal", 3,
                                                 ifelse(covsurge$Principal.Anesthesia.Technique=="Monitored Anesthesia Care/IV Sedation", 4,NA)))))

sum(is.na(covsurge$anes))
table(covsurge$anes, useNA="always") # no missing values
covsurge$anes <- as.factor(covsurge$anes)
```

## Surgery duration

```{r}
covsurge$opetime <- covsurge$Duration.of.Surgical.Procedure..in.minutes.
sum(is.na(covsurge$opetime))
```

## Transfusion

```{r}
table(covsurge$Total.Blood.Transfused..in.units.)
covsurge$transfusion <- ifelse(is.na(covsurge$Total.Blood.Transfused..in.units.), 0, covsurge$Total.Blood.Transfused..in.units.)
summary(covsurge$transfusion)
sum(is.na(covsurge$transfusion))
```

## Conversion rate

```{r}
table(covsurge$Unplanned.Conversion.to.Open)

# Conversion category: 0=No, 1=Yes
covsurge$conversion <- ifelse(covsurge$Unplanned.Conversion.to.Open=="No", 0,
                             ifelse(covsurge$Unplanned.Conversion.to.Open=="Yes", 1, NA))
covsurge$conversion[covsurge$coversion ==""] <- NA
sum(is.na(covsurge$conversion))
table(covsurge$conversion, useNA="always") 
covsurge$conversion <- as.factor(covsurge$conversion) # 3682 missing values
```

## Reoperation

```{r}
table(covsurge$First.Unplanned.return.to.OR)

# Conversion category: 0=No, 1=Yes
covsurge$reop <- ifelse(covsurge$First.Unplanned.return.to.OR=="No", 0,
                             ifelse(covsurge$First.Unplanned.return.to.OR=="Yes", 1, NA))

covsurge$reop[covsurge$coversion ==""] <- NA
sum(is.na(covsurge$reop))
table(covsurge$reop, useNA="always") # 2 missing values
covsurge$reop <- as.factor(covsurge$reop)
```

# Complication

## Thrombotic complication

```{r}
table(covsurge$X..of.Postop.Venous.Thrombosis.Requiring.Therapy)
table(covsurge$X..of.Postop.Pulmonary.Embolism)

# VTE complication category: 0=No, 1=Yes
covsurge$postop_vte <- ifelse(covsurge$X..of.Postop.Venous.Thrombosis.Requiring.Therapy == 1 |covsurge$X..of.Postop.Pulmonary.Embolism == 1, 1, 0)

table(covsurge$postop_vte)
sum(is.na(covsurge$postop_vte))
covsurge$postop_vte<- as.factor(covsurge$postop_vte)
```

## Hemorrhagic complications

```{r}
table(covsurge$X..of.Postop.Blood.Transfusions..72h.of.surgery.start.time.)

# Hemorrhagic complication category: 0=No, 1=Yes
covsurge$postop_hemo <- covsurge$X..of.Postop.Blood.Transfusions..72h.of.surgery.start.time.

sum(is.na(covsurge$postop_hemo))
covsurge$postop_hemo <- as.factor(covsurge$postop_hemo)
```

## Pulmonary complications

```{r}
table(covsurge$X..of.Postop.Pneumonia)
table(covsurge$X..of.Postop.Pneumonia.PATOS)
table(covsurge$X..of.Postop.Unplanned.Intubation)
table(covsurge$X..of.Postop.Pulmonary.Embolism)
table(covsurge$X..of.Postop.On.Ventilator...48.hours)
table(covsurge$X..of.Postop.On.Ventilator...48.hours.PATOS)

covsurge$postop_pulmo <- ifelse(covsurge$X..of.Postop.Pneumonia == 1 |covsurge$X..of.Postop.Pneumonia.PATOS == 1 |covsurge$X..of.Postop.Unplanned.Intubation >= 1 | covsurge$X..of.Postop.Pulmonary.Embolism == 1 | covsurge$X..of.Postop.On.Ventilator...48.hours >= 1 | covsurge$X..of.Postop.On.Ventilator...48.hours.PATOS == 1, 1, 0)

table(covsurge$postop_pulmo)
sum(is.na(covsurge$postop_pulmo))
covsurge$postop_pulmo <- as.factor((covsurge$postop_pulmo))
```

## Cardiac complications

```{r}
table(covsurge$X..of.Cardiac.Arrest.Requiring.CPR)
table(covsurge$X..of.Myocardial.Infarction)

covsurge$postop_cardio <- ifelse(covsurge$X..of.Cardiac.Arrest.Requiring.CPR >= 1 |covsurge$X..of.Myocardial.Infarction == 1, 1, 0)

table(covsurge$postop_cardio)
sum(is.na(covsurge$postop_cardio))
covsurge$postop_cardio <- as.factor(covsurge$postop_cardio)
```

## Neurological complications

```{r}
table(covsurge$X..of.Stroke.Cerebral.Vascular.Acccident..CVA.)
covsurge$postop_neuro <- ifelse(covsurge$X..of.Stroke.Cerebral.Vascular.Acccident..CVA. == 1, 1, 0)
table(covsurge$postop_neuro)
sum(is.na(covsurge$postop_neuro))
covsurge$postop_neuro <- as.factor(covsurge$postop_neuro)
```

## SSI complications

```{r}
table(covsurge$X..of.Postop.Superficial.Incisional.SSI)
table(covsurge$X..of.Postop.Superficial.Incisional.SSI.PATOS)
table(covsurge$X..of.Postop.Deep.Incisional.SSI)
table(covsurge$X..of.Postop.Deep.Incisional.SSI.PATOS)
table(covsurge$X..of.Postop.Organ.Space.SSI)
table(covsurge$X..of.Postop.Organ.Space.SSI.PATOS)

covsurge$ssi <- ifelse(covsurge$X..of.Postop.Superficial.Incisional.SSI >= 1 |covsurge$X..of.Postop.Superficial.Incisional.SSI.PATOS == 1 |covsurge$X..of.Postop.Deep.Incisional.SSI == 1 | covsurge$X..of.Postop.Deep.Incisional.SSI.PATOS == 1 | covsurge$X..of.Postop.Organ.Space.SSI >= 1 | covsurge$X..of.Postop.Organ.Space.SSI.PATOS == 1, 1, 0)

table(covsurge$ssi)
sum(is.na(covsurge$ssi))
covsurge$ssi <- as.factor(covsurge$ssi)
```

## Hospital stay

```{r}
covsurge$los <- covsurge$Hospital.Length.of.Stay
sum(is.na(covsurge$los))
summary(covsurge$los)
sum(is.na(covsurge$los)) # 117 missing values
```

## Death

```{r}
table(covsurge$Postop.Death.w.in.30.days.of.Procedure)

# Death category: 0=No, 1=Yes
covsurge$death <- ifelse(covsurge$Postop.Death.w.in.30.days.of.Procedure=="No", 0,
                             ifelse(covsurge$Postop.Death.w.in.30.days.of.Procedure=="Yes", 1, NA))

covsurge$death[covsurge$death ==""] <- NA
sum(is.na(covsurge$death))
table(covsurge$death, useNA="always") # 173 missing values
covsurge$death <- as.factor(covsurge$death)
sum(is.na(covsurge$death)) # 119 missing values

```

# Create new variables

```{r}
# Med staff vaccination (pre/post)
covsurge$med_vac <- as.integer(covsurge$opedate >= as.Date("2020-12-14"))
covsurge$med_vac <- as.factor(covsurge$med_vac)

# COVID-19 variant
## Delta
covsurge$delta <- as.integer(covsurge$opedate >= as.Date("2021-04-01") & covsurge$opedate <= as.Date("2022-01-31"))
covsurge$delta <- as.factor(covsurge$delta)
table(covsurge$delta)

## Omicron
covsurge$omicron <-as.integer(covsurge$opedate >= as.Date("2022-02-01")) 
covsurge$omicron <- as.factor(covsurge$omicron)
table(covsurge$omicron)

# Complication
covsurge[104:141] <- lapply(covsurge[104:141], as.numeric)
covsurge$comp <- ifelse(rowSums(covsurge[104:141], na.rm=TRUE)>0, 1, 0)
covsurge$comp <- ifelse((covsurge$comp ==0 & covsurge$death ==1)|(covsurge$comp == 1 & covsurge$death ==1), 1, covsurge$comp)
table(covsurge$comp)
# covsurge$comp <- as.factor(covsurge$comp)
```

# Missing Value

## func, dm, htn, imsp, ascites, sepsis, venti, cancer, chf, smk, copd, dialysis, aki, reop

```{r}
covsurge[is.na(covsurge$func), "IDN"]
covsurge[is.na(covsurge$dm), "IDN"]
covsurge[is.na(covsurge$htn), "IDN"]
covsurge[is.na(covsurge$imsp), "IDN"]
covsurge[is.na(covsurge$ascites), "IDN"]
covsurge[is.na(covsurge$sepsis), "IDN"]
covsurge[is.na(covsurge$venti), "IDN"]
covsurge[is.na(covsurge$cancer), "IDN"]
covsurge[is.na(covsurge$chf), "IDN"]
covsurge[is.na(covsurge$smk), "IDN"]
covsurge[is.na(covsurge$copd), "IDN"]
covsurge[is.na(covsurge$dialysis), "IDN"]
covsurge[is.na(covsurge$aki), "IDN"]
covsurge[is.na(covsurge$reop), "IDN"]
```

All of these variables are missing in the same patient. Later, I remover these two patients.

# covid

```{r}
covsurge$covid <- ifelse(covsurge$covid == 2, 1, covsurge$covid)
covsurge$covid <- as.factor(covsurge$covid)
table(covsurge$covid)
```

# Creating Final Dataset

```{r}
# At first the Dataset has 7852 patients
# Removing minor surgery patients
covsurge <- covsurge %>% filter(!(anes %in% c(0, 4))) # Then 7241 patients (= 611 patients were removed)

# Removing patients with multiple missing values
covsurge <- covsurge[covsurge$IDN !=2657842 & covsurge$IDN != 1489469,] # 2 patients were removed.

############## Finally this dataset has 7239 patients #################
```

### Main analysis

# Table 1 (Patient Demographics)

```{r}
label(covsurge$age) <- "Age"
label(covsurge$sex) <- "Sex"
label(covsurge$bmi) <- "Body Mass Index"
label(covsurge$race) <- "Race"
label(covsurge$func) <- "Functional Status"
label(covsurge$emergency) <- "Emergency Case"
label(covsurge$asa) <- "ASA-PS"
label(covsurge$dm) <- "Diabetes"
label(covsurge$htn) <- "Hypertension"
label(covsurge$imsp) <- "Steroid"
label(covsurge$ascites) <- "Ascites"
label(covsurge$sepsis) <- "Systemic sepsis"
label(covsurge$venti) <- "Ventilator"
label(covsurge$cancer) <- "Disseminated cancer"
label(covsurge$chf) <- "CHF"
label(covsurge$dyspnea) <- "Dyspnea"
label(covsurge$smk) <- "Smoker"
label(covsurge$copd) <- "COPD"
label(covsurge$dialysis) <- "Dialysis"
label(covsurge$aki) <- "AKI"
label(covsurge$med_vac) <- "Medical staff vaccine"
label(covsurge$pat_vac) <- "Patient COVID-19 vaccination"
label(covsurge$covid) <- "Preoperative Covid Infection"
label(covsurge$surg_type) <- "Type of Surgery"
label(covsurge$anes) <- "Type of Anesthesia"
label(covsurge$opetime) <- "Operation time"
label(covsurge$transfusion) <- "Transfusion"
label(covsurge$conversion) <- "Converstioin"
label(covsurge$postop_vte) <- "Thronbotic complication"
label(covsurge$postop_hemo) <- "Hemorrhagic complication"
label(covsurge$postop_pulmo) <- "Pulmonary complication"
label(covsurge$postop_cardio) <- "Cardiac complication"
label(covsurge$postop_neuro) <- "Neurological complication"
label(covsurge$ssi) <- "Surgical site complication"
label(covsurge$comp) <- "Complication"
label(covsurge$los) <- "Days in Hospital"
label(covsurge$death) <- "Death"
covsurge$surge <- factor(covsurge$surge, labels = c("Non-surge", "Surge"))

tab1 <- table1(~age+sex+bmi+race+func+emergency+asa+dm+htn+imsp+ascites+sepsis+venti+cancer+chf+dyspnea+smk+copd+dialysis+aki+med_vac+pat_vac+covid+surg_type+anes+opetime+transfusion+conversion+reop+postop_vte+postop_hemo+postop_pulmo+postop_cardio+postop_neuro+ssi+comp+los+death|surge, data=covsurge, overall = c(right="Overall"))

tab1
```

# Basic Statistics

```{r}
# Continuous variables
t.test(age ~ surge, data = covsurge)
t.test(opetime ~ surge, data = covsurge)
t.test(transfusion ~ surge, data = covsurge)
t.test(los ~ surge, data = covsurge)

# Categorical variables (chi.sq)
chisq.test(covsurge$surge, covsurge$sex)
chisq.test(covsurge$surge, covsurge$race)
chisq.test(covsurge$surge, covsurge$emergency)
chisq.test(covsurge$surge, covsurge$dm)
chisq.test(covsurge$surge, covsurge$htn)
chisq.test(covsurge$surge, covsurge$imsp)
chisq.test(covsurge$surge, covsurge$cancer)
chisq.test(covsurge$surge, covsurge$smk)
chisq.test(covsurge$surge, covsurge$copd)
chisq.test(covsurge$surge, covsurge$pat_vac)
chisq.test(covsurge$surge, covsurge$reop)
chisq.test(covsurge$surge, covsurge$postop_hemo)
chisq.test(covsurge$surge, covsurge$ssi)
chisq.test(covsurge$surge, covsurge$comp)


# Categorical variable (Fisher's exact test)
fisher.test(covsurge$surge, covsurge$bmi)
fisher.test(covsurge$surge, covsurge$func)
fisher.test(covsurge$surge, covsurge$emergency)
fisher.test(covsurge$surge, covsurge$asa)
fisher.test(covsurge$surge, covsurge$ascites)
fisher.test(covsurge$surge, covsurge$sepsis)
fisher.test(covsurge$surge, covsurge$venti)
fisher.test(covsurge$surge, covsurge$chf)
fisher.test(covsurge$surge, covsurge$dyspnea)
fisher.test(covsurge$surge, covsurge$dialysis)
fisher.test(covsurge$surge, covsurge$aki)
fisher.test(covsurge$surge, covsurge$med_vac)
fisher.test(covsurge$surge, covsurge$covid)
fisher.test(covsurge$surge, covsurge$surg_type, simulate.p.value=T)
fisher.test(covsurge$surge, covsurge$anes)
fisher.test(covsurge$surge, covsurge$conversion)
fisher.test(covsurge$surge, covsurge$postop_vte)
fisher.test(covsurge$surge, covsurge$postop_pulmo)
fisher.test(covsurge$surge, covsurge$postop_cardio)
fisher.test(covsurge$surge, covsurge$postop_neuro)
fisher.test(covsurge$surge, covsurge$death)


#chisq.test(covsurge$surge, covsurge$pat_vac)
```
# Main analysis
```{r}
model1 <- glm(comp ~ surge + age + sex + race + bmi + ses + asa + func + dm + htn + imsp + ascites + sepsis + venti + cancer + chf + copd + dialysis + aki + covid +pat_vac + delta, family=binomial(link = 'logit'), data = covsurge)
summary(model1)
exp(coef(model1))
exp(confint(model1))
```

# Secondary Analysis
```{r}
# Death
model2 <- glm(death ~ surge + age + sex + race + bmi + dm + htn + cancer + pat_vac + copd  + covid , family=binomial(link = 'logit'), data = covsurge)
summary(model2)
exp(coef(model2))
exp(confint(model2))

# Reoperation
model3 <- glm(reop ~ surge + age + sex + race + bmi + dm + htn + cancer + pat_vac + copd  + covid , family=binomial(link = 'logit'), data = covsurge)
summary(model3)
exp(coef(model3))
exp(confint(model3))

# VTE
model4 <- glm(postop_vte ~ surge + age + sex + race + bmi + dm + htn + cancer + pat_vac + copd  + covid , family=binomial(link = 'logit'), data = covsurge)
summary(model4)
exp(coef(model4))
exp(confint(model4))

# pulmonary
model5 <- glm(postop_pulmo ~ surge + age + sex + race + bmi + dm + htn + cancer + pat_vac + copd  + covid , family=binomial(link = 'logit'), data = covsurge)
summary(model5)
exp(coef(model5))
exp(confint(model5))

# Hemorrhage
model6 <- glm(postop_pulmo ~ surge + age + sex + race + bmi + dm + htn + cancer + pat_vac + copd  + covid , family=binomial(link = 'logit'), data = covsurge)
summary(model6)
exp(coef(model6))
exp(confint(model6))

```

# Test of Homogeneity
```{r}
# Omicron
table(covsurge$surge, covsurge$omicron) ## No omicron during the "surge" period

# Delta
table(covsurge$surge, covsurge$delta)

model7 <- glm(comp ~ surge + age + sex + race + bmi + ses + asa + func + dm + htn + imsp + ascites + sepsis + venti + cancer + chf + copd + dialysis + aki + covid +pat_vac + delta + surge*delta , family=binomial(link = 'logit'), data = covsurge)
summary(model7)
exp(coef(model7))
exp(confint(model7))

```


# Sensitivity Analysis
## Main Analysis with Multiple imputation
```{r}
# BMI (47 missing values)
# ASA-PS (37 missing values)
# Race (30 missing values)
multiple <- covsurge[,c("comp", "surge", "age", "sex", "race", "bmi", "ses", "asa", "func", "dm", "htn", "imsp", "ascites", "sepsis", "venti", "cancer", "chf", "copd", "dialysis", "aki", "pat_vac", "covid", "delta")]

# Check predictors (1 means that that column is used to impute the row variable; 0 means no use)
test <- mice(multiple, m=5, print=F)
test
# Select variables to be analyzed
predM <- test$predictorMatrix
predM[c("comp", "surge", "age", "sex", "ses", "func", "dm", "htn", "imsp", "ascites", "sepsis", "venti", "cancer", "chf", "copd", "dialysis", "aki", "pat_vac", "covid", "delta"),] <- 0
predM[,c("comp", "surge", "age", "sex", "ses", "func", "dm", "htn", "imsp", "ascites", "sepsis", "venti", "cancer", "chf", "copd", "dialysis", "aki", "pat_vac", "covid", "delta")] <- 0 
predM

# multi-imputation, here m=30
mi <- mice(multiple, m=20, matix=50, predictorMatrix = predM, print = F, seed = 1005)
primary_mod <- with(mi, stats::glm(comp ~ surge + age + sex + race + bmi + ses + asa + func + dm + htn + imsp + ascites + sepsis + venti + cancer + chf + copd + dialysis + aki + covid +pat_vac + delta, family=binomial(link = 'logit'), data = covsurge))

# Create complete dataset
dataset <- compliete(mi)

# primary_mod
miced_pool <- pool(primary_mod)
summary(miced_pool)
coefficients((miced_pool))
coef(miced_pool)

beta <- -0.088667469
exp(beta)
SE <- 0.1101750
lcl <- beta-qnorm(0.975)*SE 
ucl <- beta+qnorm(0.975)*SE
cbind(beta, lcl, ucl)
```

# IPWeighting for confounding and selection bias
```{r}
covsurge$cens <- ifelse(is.na(covsurge$comp), 1, 0)

table(covsurge$cens)
prop.table(table(covsurge$cens))

################################################
# Adjusting for confounding and selection bias #
################################################

# estimation of denominator of treatment weights
denom.tx <- glm(surge ~ age + sex + race + bmi + ses + asa + func + dm + htn + imsp + ascites + sepsis + venti + cancer + chf + copd + dialysis + aki + covid + pat_vac + delta, family = binomial(), data = covsurge)
summary(denom.tx)
pd.surge <- predict(denom.tx, type = "response")

# estimation of numerator of treatment weights
numer.tx <- glm(surge~1, family = binomial(), data = covsurge)
summary(numer.tx)
pn.surge <- predict(numer.tx, type = "response")

# estimation of denominator of censoring weights
denom.cens <- glm(cens ~ surge + age + sex + race + bmi + ses + asa + func + dm + htn + imsp + ascites + sepsis + venti + cancer + chf + copd + dialysis + aki + covid + pat_vac + delta, family = binomial(), data = covsurge)
summary(denom.cens)

pd.cens <- 1-predict(denom.cens, type = "response")

# estimation of numerator of censoring weights
numer.cens <- glm(cens~surge, family = binomial(), data = covsurge)
summary(numer.cens)
pn.cens <- 1-predict(numer.cens, type = "response")

covsurge$sw.a <- ifelse(covsurge$surge == 0, ((1-pn.surge)/(1-pd.surge)),
                     (pn.surge/pd.surge))
covsurge$sw.c <- pn.cens/pd.cens
covsurge$sw <- covsurge$sw.c*covsurge$sw.a

summary(covsurge$sw.a)
#sd(covsurge$sw.a)
#summary(covsurge$sw.c)
#sd(covsurge$sw.c)
summary(covsurge$sw)
#sd(covsurge$sw)

msm.sw <- geeglm(comp~surge, data=covsurge, 
                 weights=sw, id=IDN, corstr="independence")
summary(msm.sw)

beta <- coef(msm.sw)
SE <- coef(summary(msm.sw))[,2]
lcl <- beta-qnorm(0.975)*SE 
ucl <- beta+qnorm(0.975)*SE
cbind(beta, lcl, ucl)
cbind(exp(beta), exp(lcl), exp(ucl))
```

# IP weighting for selection bias and propensity score for confounding
```{r}
# estimation of denominator of censoring weights
#################### bmi, race, covid have missing value, then I removed them. ###############################
denom.cens <- glm(cens ~ surge + age + sex + race + bmi + ses + asa + func + dm + htn + imsp + ascites + sepsis + venti + cancer + chf + copd + dialysis + aki + covid + pat_vac + delta, family = binomial(), data = covsurge)
summary(denom.cens)

covsurge.c <- covsurge[which(covsurge$cens!=1),]
covsurge.c$pd.cens <- predict(denom.cens, covsurge.c, type = "response")
covsurge.c$wc <- 1/(covsurge.c$pd.cens)
summary(covsurge.c$wc)

# PS: Parametric estimation with many covariates
fit.para2 <- glm(surge ~ age + sex + race + bmi + ses + asa + func + dm + htn + imsp + ascites + sepsis + venti + cancer + chf + copd + dialysis + aki + covid + pat_vac + delta, data=covsurge.c, family=binomial())
summary(fit.para2)

covsurge.c$p.surge <- predict(fit.para2, type="response")

fit.ps <- glm(comp~surge+p.surge+I(p.surge*p.surge), data= covsurge.c,  weight=wc)
summary(fit.ps)
exp(coef(fit.ps))
exp(confint(fit.ps))
```

# Create the changes in complications
```{r}
covsurge.nmv <- covsurge[which(!is.na(covsurge$comp)),]
# Group by month and calculate total complications for each month
monthly_comp <- covsurge.nmv %>%
  mutate(month = format(opedate, "%Y-%m")) %>% # Create month columns for grouping by month
  group_by(month) %>% 
  summarise(complication = sum(comp), ope = n_distinct(IDN)) %>%
  mutate(comp_rate = complication/ope)

# Change the variable class
monthly_comp$month <- as.Date(paste0(monthly_comp$month, "-01"))
monthly_comp$ope <- as.numeric(monthly_comp$ope)

summary(monthly_comp$comp_rate) # Mean Complication rate is 0.1598

# Graph plotting
ggplot(monthly_comp, aes(x = month, y = complication)) +
  geom_line() + 
  scale_y_continuous(limits = c(0, 60)) 

ggplot(monthly_comp, aes(x = month, y = ope)) +
  geom_line()

ggplot(monthly_comp, aes(x = month, y = comp_rate)) +
  geom_line(color = "darkred", size = 1.2) +
  scale_x_date(date_labels = "%Y-%m", date_breaks = "3 months") +
  scale_y_continuous(limits = c(0, 0.3)) +
  xlab("Date") + ylab("Incidence of Complication") + 
  theme_economist() +
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5, hjust=1)) +
  geom_abline(intercept = 0.1598, slope = 0, lty = 2, color = "darkgrey") + 
  ggtitle("The Trends in Complication Incidence") + 
  geom_vline(xintercept = as.numeric(as.Date("2020-03-01")), color = "red")

```


