---
title: "covid_surge"
author: "Takuto Yoshida"
date: "2023-02-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Install package
```{r}
if (!require(readxl)) {install.packages("readxl")}
if (!require(ggplot2)) {install.packages("ggplot2")}
if (!require(dplyr)) {install.packages("dplyr")}
if (!require(car)) {install.packages("car")}
if (!require(car)) {install.packages("tidyr")}
if (!require(car)) {install.packages("stringi")}
if (!require(lubridate)) {install.packages("lubridate")}
if (!require(mice)) {install.packages("mice")}
library(lubridate)

pacman::p_load(
  rio,          # file import
  here,         # locate file
  skimr,        # Summarized data
  tidyverse,    # Data management and ggplot2 drawing
  gtsummary,    # Summary Statistics and Tests
  rstatix,      # Summary Statistics and Statistical Tests
  janitor,      # Add totals and percentages to tables
  scales,       # Convert to percentages
  flextable,     # Convert tables to clean images
  lubridate,  # 日付の扱いと変換のための一般的なパッケージ  
  parsedate,  # 厄介な日付を「推測」する関数を持つ
  aweek,      # 日付を週に、週を日付に変換する別のオプション
  zoo,        # 日付と時間に関する追加の関数
  tidyverse,  # データマネジメントと可視化  
  rio)        # データのインポートとエクスポート)
```

# Create the dataset
```{r}
covsurge <- read.csv("/Users/yoshidatakuto/Dropbox/HSPH/MPH-CLE/EPI 946/Dataset/nsqip_op.csv")
names(covsurge)
```

## Operation Day
```{r}
# Separate Operation.Date into Year, Month, and Day
class(covsurge$Operation.Date)
covsurge <- covsurge %>%
  mutate(opedate = as.Date(Operation.Date, format ="%m/%d/%Y"))
class(covsurge$opedate)

# Extract data after March 1, 2020.
covsurge <- covsurge %>%
  filter(opedate >= as.Date("2020-03-01"))
```

# COVID Surge
```{r}
# Define a date range during the surge period
## December11, 2020 - March1, 2021
## December27, 2021 - January31, 2022
surge_start_dates <- c("2020-12-11", "2021-12-27")
surge_end_dates <- c("2021-03-01", "2022-01-31")

# Assign 1 to the data corresponding to the surge period 
covsurge$surge <- 0
for (i in 1:length(surge_start_dates)) {
  covsurge$surge[covsurge$opedate >= as.Date(surge_start_dates[i]) & 
                  covsurge$opedate <= as.Date(surge_end_dates[i])] <- 1
}

# Check for correct assignment
# subset(covsurge, surge == 0 ) $opedate
# subset(covsurge, surge == 1 ) $opedate
```

## Age
```{r}
# countinuous
covsurge$age <- covsurge$Age.at.Time.of.Surgery

# Age category
covsurge$age_cat <- ifelse(covsurge$age < 40, 1, 
                           ifelse(covsurge$age <=50, 2,
                                  ifelse(covsurge$age <= 60, 3, 
                                         ifelse(covsurge$age <= 70, 4, 5))))
summary(covsurge$age)
table(covsurge$age_cat)
sum(is.na(covsurge$age)) # No missing value
```

## Sex
```{r}
covsurge$sex <- ifelse(covsurge$Sex == "Male", 1, 0) # Female:0, Male:1
covsurge$sex <- as.factor(covsurge$sex)
table(covsurge$sex)
sum(is.na(covsurge$sex)) # No missing value
```

## Race
```{r}
table(covsurge$Race)
covsurge$Race[covsurge$Race ==""] <- NA

# Race=0: Black or African American, Race=1: White, Race=2: Others
covsurge$race <- ifelse(covsurge$Race=="Black or African American", 0,
                        ifelse(covsurge$Race=="White", 1,2))

covsurge$race <- as.factor(covsurge$race)
sum(is.na(covsurge$race))
table(covsurge$race, useNA="always")  # 24 missing values
```

## BMI
```{r}
covsurge$bmi <- ifelse(covsurge$BMI < 18.5, 0,  # BMI<18.5: 0
                           ifelse(covsurge$BMI <25, 1, # BMI 18.5-25: 1
                                  ifelse(covsurge$BMI < 30, 2, # BMI 25-30: 2
                                         ifelse(covsurge$BMI >= 30, 3, NA)))) # BMI 30-: 3, missing value: NA 
table(covsurge$bmi, useNA="always")
sum(is.na(covsurge$bmi))  
covsurge$bmi <- as.factor(covsurge$bmi) # 47 missing value
```

## Functional Status
```{r}
table(covsurge$Functional.Heath.Status)
covsurge$Functional.Heath.Status[covsurge$Functional.Heath.Status ==""] <- NA

# func=0: Independent, func=1: Partially Dependent, func=2: Tottally Dependent, func=3: Unknown
covsurge$func <- ifelse(covsurge$Functional.Heath.Status=="Independent", 0,
                        ifelse(covsurge$Functional.Heath.Status=="Partially Dependent", 1, 
                        ifelse(covsurge$Functional.Heath.Status=="Totally Dependent", 2,
                        ifelse(covsurge$Functional.Heath.Status=="Unknown",3,NA))))

sum(is.na(covsurge$Functional.Heath.Status))
table(covsurge$func, useNA="always") 
covsurge$func <- as.factor(covsurge$func) # two missing value
```

## Emergency
```{r}
table(covsurge$Emergency.Case)

# emergency=0: Elective surgery, emergency=1: Emergency surgery
covsurge$emergency <- ifelse(covsurge$Emergency.Case=="No", 0,
                             ifelse(covsurge$Emergency.Case=="Yes", 1, NA))
sum(is.na(covsurge$emergency))
table(covsurge$emergency, useNA="always")
covsurge$emergency <- as.factor(covsurge$emergency) # 5612 missing value
```
More than 50% are missing values, so they are not incorporated into the final model. This will be added to Limitation.

## ASA-PS
```{r}
table(covsurge$ASA.Classification)

# ASA-PS category
covsurge$asa <- ifelse(covsurge$ASA.Classification=="ASA 1 - No Disturb", 1,
                             ifelse(covsurge$ASA.Classification=="ASA 2 - Mild Disturb", 2,
                                    ifelse(covsurge$ASA.Classification=="ASA 3 - Severe Disturb", 3,
                                           ifelse(covsurge$ASA.Classification=="ASA 4 - Life Threat", 4,
                                                  ifelse(covsurge$ASA.Classification=="ASA 5 - Moribund", 5, NA)))))
sum(is.na(covsurge$asa))
table(covsurge$asa, useNA="always") 
covsurge$asa <- as.factor(covsurge$asa) # 37 missing values
```

## DM
```{r}
table(covsurge$Diabetes.Mellitus)

# DM category: 0=No, 1=Non-Insulin, 2=Insulin
covsurge$dm <- ifelse(covsurge$Diabetes.Mellitus=="No", 0,
                             ifelse(covsurge$Diabetes.Mellitus=="Non-Insulin", 1,
                                    ifelse(covsurge$Diabetes.Mellitus=="Insulin", 2, NA)))
sum(is.na(covsurge$dm))
table(covsurge$dm, useNA="always") 
covsurge$dm <- as.factor(covsurge$dm) # 2 missing values
```

## HTN
```{r}
table(covsurge$Hypertension.requiring.medication)

# HTN category: 0=No, 1=Yes
covsurge$htn <- ifelse(covsurge$Hypertension.requiring.medication=="No", 0,
                             ifelse(covsurge$Hypertension.requiring.medication=="Yes", 1, NA))
sum(is.na(covsurge$htn))
table(covsurge$htn, useNA="always") 
covsurge$htn <- as.factor(covsurge$htn) # 2 missing values
covsurge$htn <- as.factor(covsurge$htn)
```

## Immunosuppressive Therapy
```{r}
table(covsurge$Immunosuppressive.Therapy)

# Immunosuppressive category: 0=No, 1=Yes
covsurge$imsp <- ifelse(covsurge$Immunosuppressive.Therapy=="No", 0,
                             ifelse(covsurge$Immunosuppressive.Therapy=="Yes", 1, NA))
sum(is.na(covsurge$imsp))
table(covsurge$imsp, useNA="always") 
covsurge$imsp <- as.factor(covsurge$imsp) # 2 missing values
```

## Ascites
```{r}
table(covsurge$Ascites)

# Ascites category: 0=No, 1=Yes
covsurge$ascites <- ifelse(covsurge$Ascites=="No", 0,
                             ifelse(covsurge$Ascites=="Yes", 1, NA))
sum(is.na(covsurge$ascites))
table(covsurge$ascites, useNA="always") 
covsurge$ascites <- as.factor(covsurge$ascites) # 2 missing values
```

## Systemic sepsis within 48 h before surgery 
```{r}
table(covsurge$Sepsis..SIRS.Sepsis.Septic.shock...48h.)

# Sepsis category: 0=None, 1=Sepsis, 2=Septic shock, 3=SIRS
covsurge$sepsis <- ifelse(covsurge$Sepsis..SIRS.Sepsis.Septic.shock...48h.=="None", 0,
                             ifelse(covsurge$Sepsis..SIRS.Sepsis.Septic.shock...48h.=="Sepsis", 1,
                                    ifelse(covsurge$Sepsis..SIRS.Sepsis.Septic.shock...48h.=="Septic shock", 3,
                                           ifelse(covsurge$Sepsis..SIRS.Sepsis.Septic.shock...48h.=="SIRS", 4, NA))))
sum(is.na(covsurge$sepsis))
table(covsurge$sepsis, useNA="always") 
covsurge$sepsis <- as.factor(covsurge$sepsis) # 2 missing values
```

## Ventilator dependency
```{r}
table(covsurge$Ventilator.Dependent)

# Ventilator dependency category: 0=No, 1=Yes
covsurge$venti <- ifelse(covsurge$Ventilator.Dependent=="No", 0,
                             ifelse(covsurge$Ventilator.Dependent=="Yes", 1, NA))
sum(is.na(covsurge$venti))
table(covsurge$venti, useNA="always") 
covsurge$venti <- as.factor(covsurge$venti) # 2 missing values
```

## Disseminated cancer
```{r}
table(covsurge$Disseminated.Cancer)

# Disseminated.Cancer category: 0=No, 1=Yes
covsurge$cancer <- ifelse(covsurge$Disseminated.Cancer=="No", 0,
                             ifelse(covsurge$Disseminated.Cancer=="Yes", 1, NA))
sum(is.na(covsurge$cancer))
table(covsurge$cancer, useNA="always") 
covsurge$cancer <- as.factor(covsurge$cancer) #2 missing values
```

## Congestive heart failure
```{r}
table(covsurge$Heart.Failure)

# Heart failure category: 0=No, 1=Yes
covsurge$chf <- ifelse(covsurge$Heart.Failure=="No", 0,
                             ifelse(covsurge$Heart.Failure=="Yes", 1, NA))
sum(is.na(covsurge$chf))
table(covsurge$chf, useNA="always") #2 missing values
covsurge$chf <- as.factor(covsurge$chf)
```

## Dyspnea
```{r}
table(covsurge$Dyspnea)

# Dyspnea category: 0=No, 1=Moderate Exertion, 2=AtRest
covsurge$dyspnea <- ifelse(covsurge$Dyspnea=="No", 0,
                             ifelse(covsurge$Dyspnea=="Moderate Exertion", 1,
                                    ifelse(covsurge$Dyspnea=="AtRest", 2, NA)))
covsurge$dyspnea[covsurge$dyspnea =="NA"] <- NA
sum(is.na(covsurge$dyspnea))
table(covsurge$dyspnea, useNA="always") 
covsurge$dyspnea <- as.factor(covsurge$dyspnea) #5620 missing values
```
This is also missing in about half of the cases. However, since this one is not incorporated into the final model, no assignment of missing values is made.

## Current smoker within 1y
```{r}
table(covsurge$Current.Smoker.within.1.year)

# smoking category: 0=No, 1=Yes
covsurge$smk <- ifelse(covsurge$Current.Smoker.within.1.year=="No", 0,
                             ifelse(covsurge$Current.Smoker.within.1.year=="Yes", 1, NA))
sum(is.na(covsurge$smk))
table(covsurge$smk, useNA="always") #2 missing values
covsurge$smk <- as.factor(covsurge$smk)
```

## COPD
```{r}
table(covsurge$History.of.Severe.COPD)

# COPD category: 0=No, 1=Yes
covsurge$copd <- ifelse(covsurge$History.of.Severe.COPD=="No", 0,
                             ifelse(covsurge$History.of.Severe.COPD=="Yes", 1, NA))
sum(is.na(covsurge$copd))
table(covsurge$copd, useNA="always") #2 missing values
covsurge$copd <- as.factor(covsurge$copd)
```

## Dialysis
```{r}
table(covsurge$Preop.Dialysis)

# Dialysis category: 0=No, 1=Yes
covsurge$dialysis <- ifelse(covsurge$Preop.Dialysis=="No", 0,
                             ifelse(covsurge$Preop.Dialysis=="Yes", 1, NA))
sum(is.na(covsurge$dialysis))
table(covsurge$dialysis, useNA="always") #2 missing values
covsurge$dialysis <- as.factor(covsurge$dialysis)
```

## Acute renal failure 
```{r}
table(covsurge$Preop.Acute.Kidney.Injury)

# AKI category: 0=No, 1=Yes
covsurge$aki <- ifelse(covsurge$Preop.Acute.Kidney.Injury=="No", 0,
                             ifelse(covsurge$Preop.Acute.Kidney.Injury=="Yes", 1, NA))
sum(is.na(covsurge$aki))
table(covsurge$aki, useNA="always") #2 missing values
covsurge$aki <- as.factor(covsurge$aki)
```

## COVID-19 infection at the time of surgery
```{r}
table(covsurge$Preop.COVID.19.Diagnosis)

# Preop covid category: 0=No, 1=Yes, lab-confirmed diagnosis (or ICD-10 code U07.1), 2=Yes, suspected diagnosis (or ICD-10 code U07.2)
covsurge$covid <- ifelse(covsurge$Preop.COVID.19.Diagnosis=="No", 0,
                             ifelse(covsurge$Preop.COVID.19.Diagnosis=="Yes, lab-confirmed diagnosis (or ICD-10 code U07.1)", 1,
                                    ifelse(covsurge$Preop.COVID.19.Diagnosis=="Yes, suspected diagnosis (or ICD-10 code U07.2)", 2, NA)))

covsurge$covid[covsurge$covid ==""] <- NA
sum(is.na(covsurge$covid))
table(covsurge$covid, useNA="always") 
covsurge$covid <- as.factor(covsurge$covid) # 3 missing values
```

## COVID-19 Vaccination
```{r}

```


# Operation
## Type of surgery
```{r}
table(covsurge$Surgical.Specialty)

# Type of surgery category: 0=General Surgery, 1=Neurosurgery, 2=Otolaryngology (ENT), 3=Plastics, 4=Thoracic, 5=Urology, 6=Vascular
covsurge$surg_type <- ifelse(covsurge$Surgical.Specialty=="General Surgery", 0,
                             ifelse(covsurge$Surgical.Specialty=="Neurosurgery", 1,
                                    ifelse(covsurge$Surgical.Specialty=="Otolaryngology (ENT)", 2,
                                           ifelse(covsurge$Surgical.Specialty=="Plastics", 3,
                                                 ifelse(covsurge$Surgical.Specialty=="Thoracic", 4,
                                                        ifelse(covsurge$Surgical.Specialty=="Urology", 5,
                                                               ifelse(covsurge$Surgical.Specialty=="Vascular", 6,NA)))))))

sum(is.na(covsurge$surg_type))
table(covsurge$surg_type, useNA="always") # no missing values
covsurge$surg_type <- as.factor(covsurge$surg_type)
```

## Type of Anesthesia 
```{r}
table(covsurge$Principal.Anesthesia.Technique)

# Type of surgery category: 0=No, 1=Yes, lab-confirmed diagnosis (or ICD-10 code U07.1), 2=Yes, suspected diagnosis (or ICD-10 code U07.2)
covsurge$anes <- ifelse(covsurge$Principal.Anesthesia.Technique=="Regional", 0,
                             ifelse(covsurge$Principal.Anesthesia.Technique=="General", 1,
                                    ifelse(covsurge$Principal.Anesthesia.Technique=="Epidural", 2,
                                           ifelse(covsurge$Principal.Anesthesia.Technique=="Spinal", 3,
                                                 ifelse(covsurge$Principal.Anesthesia.Technique=="Monitored Anesthesia Care/IV Sedation", 4,NA)))))

sum(is.na(covsurge$anes))
table(covsurge$anes, useNA="always") # no missing values
covsurge$anes <- as.factor(covsurge$anes)
```

## Surgery duration
```{r}
covsurge$opetime <- covsurge$Duration.of.Surgical.Procedure..in.minutes.
sum(is.na(covsurge$opetime))
```

## Transfusion
```{r}
table(covsurge$Total.Blood.Transfused..in.units.)
covsurge$transfusion <- ifelse(is.na(covsurge$Total.Blood.Transfused..in.units.), 0, covsurge$Total.Blood.Transfused..in.units.)
summary(covsurge$transfusion)
sum(is.na(covsurge$transfusion))
```

## Conversion rate
```{r}
table(covsurge$Unplanned.Conversion.to.Open)

# Conversion category: 0=No, 1=Yes
covsurge$conversion <- ifelse(covsurge$Unplanned.Conversion.to.Open=="No", 0,
                             ifelse(covsurge$Unplanned.Conversion.to.Open=="Yes", 1, NA))
covsurge$conversion[covsurge$coversion ==""] <- NA
sum(is.na(covsurge$conversion))
table(covsurge$conversion, useNA="always") 
covsurge$conversion <- as.factor(covsurge$conversion) # 3682 missing values
```

## Reoperation
```{r}
table(covsurge$First.Unplanned.return.to.OR)

# Conversion category: 0=No, 1=Yes
covsurge$reop <- ifelse(covsurge$First.Unplanned.return.to.OR=="No", 0,
                             ifelse(covsurge$First.Unplanned.return.to.OR=="Yes", 1, NA))

covsurge$reop[covsurge$coversion ==""] <- NA
sum(is.na(covsurge$reop))
table(covsurge$reop, useNA="always") # 2 missing values
covsurge$reop <- as.factor(covsurge$reop)
```

## Blood loss
```{r}

```


# Complication
## Thrombotic complication
```{r}
table(covsurge$X..of.Postop.Venous.Thrombosis.Requiring.Therapy)
table(covsurge$X..of.Postop.Pulmonary.Embolism)

# VTE complication category: 0=No, 1=Yes
covsurge$postop_vte <- ifelse(covsurge$X..of.Postop.Venous.Thrombosis.Requiring.Therapy == 1 |covsurge$X..of.Postop.Pulmonary.Embolism == 1, 1, 0)

table(covsurge$postop_vte)
sum(is.na(covsurge$postop_vte))
covsurge$postop_vte<- as.factor(covsurge$postop_vte)
```

## Hemorrhagic complications
```{r}
table(covsurge$X..of.Postop.Blood.Transfusions..72h.of.surgery.start.time.)

# Hemorrhagic complication category: 0=No, 1=Yes
covsurge$postop_hemo <- covsurge$X..of.Postop.Blood.Transfusions..72h.of.surgery.start.time.

sum(is.na(covsurge$postop_hemo))
covsurge$postop_hemo <- as.factor(covsurge$postop_hemo)
```

## Pulmonary complications
```{r}
table(covsurge$X..of.Postop.Pneumonia)
table(covsurge$X..of.Postop.Pneumonia.PATOS)
table(covsurge$X..of.Postop.Unplanned.Intubation)
table(covsurge$X..of.Postop.Pulmonary.Embolism)
table(covsurge$X..of.Postop.On.Ventilator...48.hours)
table(covsurge$X..of.Postop.On.Ventilator...48.hours.PATOS)

covsurge$postop_pulmo <- ifelse(covsurge$X..of.Postop.Pneumonia == 1 |covsurge$X..of.Postop.Pneumonia.PATOS == 1 |covsurge$X..of.Postop.Unplanned.Intubation >= 1 | covsurge$X..of.Postop.Pulmonary.Embolism == 1 | covsurge$X..of.Postop.On.Ventilator...48.hours >= 1 | covsurge$X..of.Postop.On.Ventilator...48.hours.PATOS == 1, 1, 0)

table(covsurge$postop_pulmo)
sum(is.na(covsurge$postop_pulmo))
covsurge$postop_pulmo <- as.factor((covsurge$postop_pulmo))
```

## Cardiac complications
```{r}
table(covsurge$X..of.Cardiac.Arrest.Requiring.CPR)
table(covsurge$X..of.Myocardial.Infarction)

covsurge$postop_cardio <- ifelse(covsurge$X..of.Cardiac.Arrest.Requiring.CPR >= 1 |covsurge$X..of.Myocardial.Infarction == 1, 1, 0)

table(covsurge$postop_cardio)
sum(is.na(covsurge$postop_cardio))
covsurge$postop_cardio <- as.factor(covsurge$postop_cardio)
```

## Neurological complications
```{r}
table(covsurge$X..of.Stroke.Cerebral.Vascular.Acccident..CVA.)
covsurge$postop_neuro <- ifelse(covsurge$X..of.Stroke.Cerebral.Vascular.Acccident..CVA. == 1, 1, 0)
table(covsurge$postop_neuro)
sum(is.na(covsurge$postop_neuro))
covsurge$postop_neuro <- as.factor(covsurge$postop_neuro)
```

## SSI complications
```{r}
table(covsurge$X..of.Postop.Superficial.Incisional.SSI)
table(covsurge$X..of.Postop.Superficial.Incisional.SSI.PATOS)
table(covsurge$X..of.Postop.Deep.Incisional.SSI)
table(covsurge$X..of.Postop.Deep.Incisional.SSI.PATOS)
table(covsurge$X..of.Postop.Organ.Space.SSI)
table(covsurge$X..of.Postop.Organ.Space.SSI.PATOS)

covsurge$ssi <- ifelse(covsurge$X..of.Postop.Superficial.Incisional.SSI >= 1 |covsurge$X..of.Postop.Superficial.Incisional.SSI.PATOS == 1 |covsurge$X..of.Postop.Deep.Incisional.SSI == 1 | covsurge$X..of.Postop.Deep.Incisional.SSI.PATOS == 1 | covsurge$X..of.Postop.Organ.Space.SSI >= 1 | covsurge$X..of.Postop.Organ.Space.SSI.PATOS == 1, 1, 0)

table(covsurge$ssi)
sum(is.na(covsurge$ssi))
covsurge$ssi <- as.factor(covsurge$ssi)
```

## Hospital stay
```{r}
covsurge$los <- covsurge$Hospital.Length.of.Stay
sum(is.na(covsurge$los))
summary(covsurge$los)
sum(is.na(covsurge$los)) # 117 missing values
```

## Death
```{r}
table(covsurge$Postop.Death.w.in.30.days.of.Procedure)

# Death category: 0=No, 1=Yes
covsurge$death <- ifelse(covsurge$Postop.Death.w.in.30.days.of.Procedure=="No", 0,
                             ifelse(covsurge$Postop.Death.w.in.30.days.of.Procedure=="Yes", 1, NA))

covsurge$death[covsurge$death ==""] <- NA
sum(is.na(covsurge$death))
table(covsurge$death, useNA="always") # 173 missing values
covsurge$death <- as.factor(covsurge$death)
sum(is.na(covsurge$death)) # 119 missing values

```

# Create new variables
```{r}
# Med staff vaccination (pre/post)
covsurge$med_vac <- as.integer(covsurge$opedate >= as.Date("2020-12-14"))

# COVID-19 variant
## Delta
covsurge$delta <- as.integer(covsurge$opedate >= as.Date("2021-04-01") & covsurge$opedate <= as.Date("2022-01-31"))
## Omicron
covsurge$omicron <-as.integer(covsurge$opedate >= as.Date("2022-02-01")) 
```


# Missing Value
## func, dm, htn, imsp, ascites, sepsis, venti, cancer, chf, smk, copd, dialysis, aki, reop
```{r}
covsurge[is.na(covsurge$func), "IDN"]
covsurge[is.na(covsurge$dm), "IDN"]
covsurge[is.na(covsurge$htn), "IDN"]
covsurge[is.na(covsurge$imsp), "IDN"]
covsurge[is.na(covsurge$ascites), "IDN"]
covsurge[is.na(covsurge$sepsis), "IDN"]
covsurge[is.na(covsurge$venti), "IDN"]
covsurge[is.na(covsurge$cancer), "IDN"]
covsurge[is.na(covsurge$chf), "IDN"]
covsurge[is.na(covsurge$smk), "IDN"]
covsurge[is.na(covsurge$copd), "IDN"]
covsurge[is.na(covsurge$dialysis), "IDN"]
covsurge[is.na(covsurge$aki), "IDN"]
covsurge[is.na(covsurge$reop), "IDN"]
```
All of these variables are missing in the same patient. Later, I remover these two patients.

# covid
```{r}
covsurge[is.na(covsurge$covid), "IDN"]
covsurge$covid[covsurge$covid == 2] <- 1
covsurge$covid[is.na(covsurge$covid)] <- 0

covsurge <- covsurge %>% filter(covid %in% c(0,1))
summary(covsurge$covid)
```


## Multiple imputation
```{r}
# BMI (47 missing values)
# ASA-PS (37 missing values)
# Race (30 missing values)
missing_vars <- c("bmi", "asa", "race")
covsurge_missing <- covsurge[, c("IDN", missing_vars)]

# Multiple imputation setup
impute <- mice(covsurge_missing, m = 5, maxit = 50, seed = 123)

# Retrieve assigned data
covsurge_imputed <- complete(impute)

# Combine with original data frame
covsurge <- merge(covsurge, covsurge_imputed, by = "IDN")

# check that the imputation is correct.
sum(is.na(covsurge$bmi.y))
sum(is.na(covsurge$asa.y))
sum(is.na(covsurge$race.y))
```


# Creating Final Dataset
```{r}
# At first the Dataset has 9568 patients
# Removing minor surgery patients
covsurge <- covsurge %>% filter(!(anes %in% c(0, 4))) # Then 8692 patients (= 876 patients were removed)

# Removing patients with multiple missing values
covsurge <- covsurge[covsurge$IDN !=2657842 & covsurge$IDN != 1489469,] # 2 patients were removed.

############## Finally this dataset has 8690 patients #################
```


### Main analysis ###

# Table 1 (Patient Demographics)
```{r}
covsurge %>% 
  select(age, age_cat, sex, bmi.y, race.y, func, emergency, asa.y, dm, htn, imsp, ascites, sepsis, venti, cancer, chf, dyspnea, smk, copd, dialysis, aki, covid, surg_type, anes, opetime, transfusion, conversion, reop, postop_vte, postop_hemo, postop_pulmo, postop_cardio, postop_neuro, ssi, los, death, surge) %>% 
  tbl_summary(     
    by = surge,                                               # 転帰ごとにテーブル全体を層別化
    statistic = list(all_continuous() ~ "{mean} ({sd})",        # 連続列に対して統計量を算出してフォーマット
                     all_categorical() ~ "{n} / {N} ({p}%)"),   # カテゴリ列に対して統計量を算出してフォーマット
    digits = all_continuous() ~ 2,                              # 連続列に対して丸めの指定
    type   = all_categorical() ~ "categorical",                 # 強制的に全カテゴリ水準を表示
    label  = list(                                              # 列名のラベルを表示
      surge   ~ "Surge",
      age ~ "Age(years)",
      age_cat ~ "Age category",
      sex ~ "Sex",
      bmi.y ~ "BMI",
      race.y ~ "Race",
      func ~ "Functional Status",
      emergency ~ "Emergency Surgery",
      asa.y ~ "ASA-PS",
      dm ~ "DM",
      htn ~ "HT",
      imsp ~ "Immunosuppression",
      ascites ~ "Ascites",
      sepsis ~ "Sepsis",
      venti ~ "Ventilator",
      cancer ~ "Cancer",
      chf ~ "CHF",
      dyspnea ~ "Dyspnea",
      smk ~ "Smoking",
      copd ~ "COPD",
      dialysis ~ "Dialysis",
      aki ~ "AKI",
      covid ~ "Preoperative covid infection",
      surg_type ~ "Type of Surgery",
      anes ~ "Type of Anesthesia",
      opetime ~ "Operation time",
      transfusion ~ "Transfusion",
      conversion ~ "Conversion",
      reop ~ "Reoperation",
      postop_vte ~ "VTE",
      postop_hemo ~ "bleeding",
      postop_pulmo ~ "pulmo comp",
      postop_cardio ~ "cardio comp",
      postop_neuro ~ "neuro comp",
      ssi ~ "SSI",
      los ~ "LOS",
      death ~ "Death",
    missing_text = "Missing"                                    # 欠測値の表示方法
  ))
```

# Basic Statistics
```{r}
# Continuous variables
t.test(age ~ surge, data = covsurge)
t.test(opetime ~ surge, data = covsurge)
t.test(transfusion ~ surge, data = covsurge)
t.test(los ~ surge, data = covsurge)

# Categorical variables (chi.sq)
chisq.test(covsurge$surge, covsurge$sex)
chisq.test(covsurge$surge, covsurge$race.y)
chisq.test(covsurge$surge, covsurge$func)
chisq.test(covsurge$surge, covsurge$emergency)
chisq.test(covsurge$surge, covsurge$dm)
chisq.test(covsurge$surge, covsurge$htn)
chisq.test(covsurge$surge, covsurge$imsp)
chisq.test(covsurge$surge, covsurge$cancer)
chisq.test(covsurge$surge, covsurge$smk)
chisq.test(covsurge$surge, covsurge$copd)
chisq.test(covsurge$surge, covsurge$reop)
chisq.test(covsurge$surge, covsurge$postop_hemo)
chisq.test(covsurge$surge, covsurge$ssi)


# Categorical variable (Fisher's exact test)
fisher.test(covsurge$surge, covsurge$bmi.y)
fisher.test(covsurge$surge, covsurge$func)
fisher.test(covsurge$surge, covsurge$emergency)
fisher.test(covsurge$surge, covsurge$asa.y)
fisher.test(covsurge$surge, covsurge$ascites)
fisher.test(covsurge$surge, covsurge$sepsis)
fisher.test(covsurge$surge, covsurge$venti)
fisher.test(covsurge$surge, covsurge$chf)
fisher.test(covsurge$surge, covsurge$dyspnea)
fisher.test(covsurge$surge, covsurge$dialysis)
fisher.test(covsurge$surge, covsurge$aki)
fisher.test(covsurge$surge, covsurge$med_vac)
fisher.test(covsurge$surge, covsurge$covid)
fisher.test(covsurge$surge, covsurge$surg_type, simulate.p.value=T)
fisher.test(covsurge$surge, covsurge$anes)
fisher.test(covsurge$surge, covsurge$conversion)
fisher.test(covsurge$surge, covsurge$postop_vte)
fisher.test(covsurge$surge, covsurge$postop_pulmo)
fisher.test(covsurge$surge, covsurge$postop_cardio)
fisher.test(covsurge$surge, covsurge$postop_neuro)
fisher.test(covsurge$surge, covsurge$death)


#chisq.test(covsurge$surge, covsurge$pat_vac)


```

# Main analysis
```{r}
model1 <- glm(death ~ surge + age + sex + race.y + bmi.y + dm + htn + imsp + sepsis + venti + cancer + chf + copd + dialysis + aki + med_vac + anes + covid + omicron + delta, family=binomial(link = 'logit'), data = covsurge)
summary(model1)
exp(coef(model1))
exp(confint(model1))
```

# Secondary Analysis
```{r}
# VTE
model2 <- glm(postop_vte ~ surge + age + sex + race.y + bmi.y + dm + htn + imsp + sepsis + venti + cancer + chf + copd + dialysis + aki + med_vac + anes + covid + omicron + delta, family=binomial(link = 'logit'), data = covsurge)
summary(model2)
exp(coef(model2))
exp(confint(model2))

# Bleeding
model3 <- glm(postop_hemo ~ surge + age + sex + race.y + bmi.y + dm + htn + imsp + sepsis + venti + cancer + chf + copd + dialysis + aki + med_vac + anes + covid + omicron + delta, family=binomial(link = 'logit'), data = covsurge)
summary(model3)
exp(coef(model3))
exp(confint(model3))

# pulmonary
model4 <- glm(postop_pulmo ~ surge + age + sex + race.y + bmi.y + dm + htn + imsp + sepsis + venti + cancer + chf + copd + dialysis + aki + med_vac + anes + covid + omicron + delta, family=binomial(link = 'logit'), data = covsurge)
summary(model4)
exp(coef(model4))
exp(confint(model4))

# cardiac
model5 <- glm(postop_cardio ~ surge + age + sex + race.y + bmi.y + dm + htn + imsp + sepsis + venti + cancer + chf + copd + dialysis + aki + med_vac + anes + covid + omicron + delta, family=binomial(link = 'logit'), data = covsurge)
summary(model5)
exp(coef(model5))
exp(confint(model5))

# ssi
model6 <- glm(ssi ~ surge + age + sex + race.y + bmi.y + dm + htn + imsp + sepsis + venti + cancer + chf + copd + dialysis + aki + med_vac + anes + covid + omicron + delta, family=binomial(link = 'logit'), data = covsurge)
summary(model6)
exp(coef(model6))
exp(confint(model6))

# neuro
model7 <- glm(postop_neuro ~ surge + age + sex + race.y + bmi.y + dm + htn + imsp + sepsis + venti + cancer + chf + copd + dialysis + aki + med_vac + anes + covid + omicron + delta, family=binomial(link = 'logit'), data = covsurge)
summary(model7)
exp(coef(model7))
exp(confint(model7))
```

# Test of Homogeneity
```{r}

```

# Sensitivity Analysis
```{r}

```





















