---
title: "Nurse staffing"
author: "Takuto Yoshida"
date: "2023-01-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install packages
```{r}
install.packages("readxl")

```

## Library
```{r}
library(readxl) # importing excel file
library(ggplot2) # creating figure
library(dplyr) # data manupulation

pacman::p_load(
  rio,          # file import
  here,         # locate file
  skimr,        # Summarized data
  tidyverse,    # Data management and ggplot2 drawing
  gtsummary,    # Summary Statistics and Tests
  rstatix,      # Summary Statistics and Statistical Tests
  janitor,      # Add totals and percentages to tables
  scales,       # Convert to percentages
  flextable     # Convert tables to clean images
  )
```

## Importing Data
```{r}
df_or <- read_excel("/Users/yoshidatakuto/Dropbox/HSPH/MPH-CLE/EPI 946/Dataset/Staffing_Data.xlsx")

# nsqip <- read.csv("")

```

## Checking the data
```{r}
# OR data
str(df_or)
summary(df_or)
skim(df_or)
head(df_or)

# NSQIP data
# str(nsqip)
# summary(nsqip)
# skim(nsqip)
# head(nsqip)
```
There are missing values in the Traveler and LOA categories, but those need to be checked.

## Dealing with missing values
Cleaning data by assuming all missing values are 0.
```{r}
df_or[is.na(df_or)] <- 0
```


## Creating new variables
```{r}
# Current FTE
df_or$Current_FTE <- df_or$Budget - df_or$Vacancy - df_or$LOA + df_or$Traveler

# TN rate (Traveler Nurse rate)
df_or$TN_rate <- df_or$Traveler / df_or$Current_FTE
```

## Figure1
```{r}
# Filter data for OR - West group
df_or_west <- df_or %>% filter(`Cost Center Name`=="OR - West")
summary(df_or_west)

# Add a new variable "period" that represents the period of 2018.1 - 2022.12
df_or_west$period <- (df_or_west$Year-2018)*12 + df_or_west$Month

# Plot the graph between Current_FTE and Travelor nurse
ggplot(df_or_west, aes(x = period, y = Traveler, color = "Traveler")) +
  geom_line()+
  geom_line(aes(x = period, y = Current_FTE, color = "Current_FTE")) +
  scale_x_continuous(limits = c(1, 71), breaks = seq(1, 72, 12), labels = c("2018.1","2018.12","2019.12","2020.12","2021.12","2022.12")) +
  labs(x = "Period (2018.1-2022.12)", y="Number of Nurse", color="Variable") +
  theme_classic()

# Plot the graph of TN_rate
ggplot(df_or_west, aes(x = period, y = TN_rate, color = "TN rate")) +
  geom_line() +
  scale_x_continuous(limits = c(1, 71), breaks = seq(1, 72, 12), labels = c("2018.1","2018.12","2019.12","2020.12","2021.12","2022.12"))  +
  labs(x = "Period (2018.1-2022.12)", y="Value", color="Variable") +
  theme_classic()

```

```{r}
# Prepandemic dataframe from 2018.1 - 2020.12
df_or_west_prepandemic <- df_or_west[df_or_west$Year >= 2018 & df_or_west$Year <= 2020,]

table(df_or_west_prepandemic$Year, df_or_west_prepandemic$Month)
table(df_or$Year, df_or$Month)

# Baseline Current_FTE and TN_rate (from2018.1 - 2020.12)
## Mean
mean(df_or_west_prepandemic$Current_FTE)
mean(df_or_west_prepandemic$TN_rate)

## Man of TN_rate (from 2018.1 - 2020.12)
max(df_or_west_prepandemic$TN_rate)

## Min of Current_FTE (from 2018.1 - 2020.12)
min(df_or_west_prepandemic$Current_FTE)
```

```{r}
# Extract TN_rate >= 0.064 from "df_or_west"
df_high_tn_rate <- df_or_west[df_or_west$TN_rate >= 0.064, c("Year", "Month")]
print(df_high_tn_rate) # All periods after April 2021 are equivalent.

# Extract Current_FTE < 34.0
df_low_curfte <- df_or_west[df_or_west$Current_FTE < 34.0, c("Year", "Month")]
print(df_low_curfte) # July 2021, and Nov 2021 or later are equivalent.

# Create variable of high_tn_rate and quant_short in nsqip
nsqip$high_tn_rate <- ifelse(nsqip$Year >= 2021 & nsqip$Month >= 4, 1, 0)
nsqip$quant_short <- ifelse(nsqip$Year = 2021 & nsqip$Month = 7, 1, 
                            ifelse(nsqip$Year >= 2021 & nsqip$Month >= 11, 1, 0))

```

## pat_vaccination
## med_staff_vaccination
## cov_surge

## summary statistics

## log transformation
## check the assumption of linear regression

## if it fail, we will do medican regression

## EMM(covid variant) check

# Subgroup analysis
## Create the data for Delta and Omicron variant
## log transformation or median regression

























