---
title: "Nurse staffing"
author: "Takuto Yoshida"
date: "2023-01-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install packages
```{r}
install.packages("readxl")

```

## Library
```{r}
library(readxl) # importing excel file
library(ggplot2) # creating figure
library(dplyr) # data manupulation
library(car) # check VIF

pacman::p_load(
  rio,          # file import
  here,         # locate file
  skimr,        # Summarized data
  tidyverse,    # Data management and ggplot2 drawing
  gtsummary,    # Summary Statistics and Tests
  rstatix,      # Summary Statistics and Statistical Tests
  janitor,      # Add totals and percentages to tables
  scales,       # Convert to percentages
  flextable     # Convert tables to clean images
  )
```

## Importing Data
```{r}
df_or <- read_excel("/Users/yoshidatakuto/Dropbox/HSPH/MPH-CLE/EPI 946/Dataset/Staffing_Data.xlsx")

# nsqip <- read.csv("")

```

## Checking the data
```{r}
# OR data
str(df_or)
summary(df_or)
skim(df_or)
head(df_or)

# NSQIP data
# str(nsqip)
# summary(nsqip)
# skim(nsqip)
# head(nsqip)
```
There are missing values in the Traveler and LOA categories, but those need to be checked.

## Dealing with missing values
Cleaning data by assuming all missing values are 0.
```{r}
df_or[is.na(df_or)] <- 0
```


## Creating new variables
```{r}
# Current FTE
df_or$Current_FTE <- df_or$Budget - df_or$Vacancy - df_or$LOA + df_or$Traveler

# TN rate (Traveler Nurse rate)
df_or$TN_rate <- df_or$Traveler / df_or$Current_FTE
```

## Figure1
```{r}
# Filter data for OR - West group
df_or_west <- df_or %>% filter(`Cost Center Name`=="OR - West")
summary(df_or_west)

# Add a new variable "period" that represents the period of 2018.1 - 2022.12
df_or_west$period <- (df_or_west$Year-2018)*12 + df_or_west$Month

# Plot the graph between Current_FTE and Travelor nurse
ggplot(df_or_west, aes(x = period, y = Traveler, color = "Traveler")) +
  geom_line()+
  geom_line(aes(x = period, y = Current_FTE, color = "Current_FTE")) +
  scale_x_continuous(limits = c(1, 71), breaks = seq(1, 72, 12), labels = c("2018.1","2018.12","2019.12","2020.12","2021.12","2022.12")) +
  labs(x = "Period (2018.1-2022.12)", y="Number of Nurse", color="Variable") +
  theme_classic()

# Plot the graph of TN_rate
ggplot(df_or_west, aes(x = period, y = TN_rate, color = "TN rate")) +
  geom_line() +
  scale_x_continuous(limits = c(1, 71), breaks = seq(1, 72, 12), labels = c("2018.1","2018.12","2019.12","2020.12","2021.12","2022.12"))  +
  labs(x = "Period (2018.1-2022.12)", y="Value", color="Variable") +
  theme_classic()

```

```{r}
# Prepandemic dataframe from 2018.1 - 2020.12
df_or_west_prepandemic <- df_or_west[df_or_west$Year >= 2018 & df_or_west$Year <= 2020,]

table(df_or_west_prepandemic$Year, df_or_west_prepandemic$Month)
table(df_or$Year, df_or$Month)

# Baseline Current_FTE and TN_rate (from2018.1 - 2020.12)
## Mean
mean(df_or_west_prepandemic$Current_FTE)
mean(df_or_west_prepandemic$TN_rate)

## Man of TN_rate (from 2018.1 - 2020.12)
max(df_or_west_prepandemic$TN_rate)

## Min of Current_FTE (from 2018.1 - 2020.12)
min(df_or_west_prepandemic$Current_FTE)
```

```{r}
# Extract TN_rate >= 0.064 from "df_or_west"
df_high_tn_rate <- df_or_west[df_or_west$TN_rate >= 0.064, c("Year", "Month")]
print(df_high_tn_rate) # All periods after April 2021 are equivalent.

# Extract Current_FTE < 34.0
df_low_curfte <- df_or_west[df_or_west$Current_FTE < 34.0, c("Year", "Month")]
print(df_low_curfte) # July 2021, and Nov 2021 or later are equivalent.

# Create variable of high_tn_rate and quant_short in nsqip
# nsqip$high_tn_rate <- ifelse(nsqip$Year >= 2021 & nsqip$Month >= 4, 1, 0)
# nsqip$quant_short <- ifelse(nsqip$Year = 2021 & nsqip$Month = 7, 1, 
#                             ifelse(nsqip$Year >= 2021 & nsqip$Month >= 11, 1, 0))

```

## Create new variables
```{r}
# patient vaccination was started on Dec 2020.
# nsqip$pat_vac <- ifelse(nsqip$Year >= 2020 & nsqip$Month >= 12, 1, 0)

# Med staff vaccination was started on Dec, 2020 as well
# nsqip$med_staff_vac <- ifelse(nsqip$Year >= 2020 & nsqip$Month >= 12, 1, 0)

# COVID-19 was surged in March20,2020- June20,2020, Oct16,2020-May31,2021, July20,2021-March15,2022
# cov_surg <- ifelse(nsqip$Year == 2020 & nsqip$Month >= 3 & nsqip$)
# nsqip$cov_surg[which(nsqip$Year == 2020 & (nsqip$Month == 3 & nsqip$Date >= 20) | 
#                      (nsqip$Month >= 4 & nsqip$Month <= 6) | 
#                      (nsqip$Month == 6 & nsqip$Date <= 20))] <- 1
# nsqip$cov_surg[which(nsqip$Year == 2020 & nsqip$Month == 10 & nsqip$Date >= 16) |
#                       (nsqip$Year == 2021 &nsqip$Month >= 11) |
#                       (nsqip$Year == 2021 & nsqip$Month >= 1 & nsqip$ Month <= 5)] <- 1
# nsqip$cov_surg[which(nsqip$Year == 2021 & nsqip$Month >= 7 & nsqip$Day >= 20) |
#                  (nsqip$Year == 2021 & nsqip$Month >= 8) |
#                  (nsqip$Year == 2022 & nsqip$Month <=4) |
#                  (nsqip$Year == 2022 & nsqip$Month == 5 & nsqip$Day <= 15)] <- 1
```

# Table1
## summary statistics
```{r}
linelist %>% 
  select(age_at_time_of_surgery, sex, race, bmi, dm, cancer, ckd, ht, copd, pat_vac, cov_infec, elective_surgery, ) %>% # variables of interest
  tbl_summary(     
    by = high_tn_rate,                                          # group by high_tn_rate
    statistic = list(all_continuous() ~ "{mean} ({sd})",        # for continuous variable
                     all_categorical() ~ "{n} / {N} ({p}%)"),   # for categorical variable
    digits = all_continuous() ~ 1,                              # round the numbers
    type   = all_categorical() ~ "categorical",                 # Forced display of all category levels
    label  = list(                                              # Display column name labels
      age_at_time_of_surgery ~ "Age",
      sex   ~ "Sex",                           
      race ~ "Race",
      bmi    ~ "BMI",
      dm ~ "Diabetes",
      cancer ~ "Cancer",
      ckd ~ "Chronic Kidney Disease",
      ht ~ "Hypertension",
      copd ~ "COPD",
      pat_vac ~ "Patient Vaccination at the time of surgery",
      cov_infec ~ "Covid-19 infection at the time of surgery",
      elective_surgery ~ "Type of surgery"
      cci ~ "Comprehensive, Complication Index",
      missing_text = "Missing"                                    # Display missing value
  ))

### Uncertain variables that can be extracted
# Socioeconomic Status, CCI, Surgical difficulty, Clavien-Dindo
 
```

# Figure 2
## Line graph showing when surgeries were performed during periods with nursing shortages or not.
```{r}
ggplot(nsqip, aes(x = as.Date(paste(Year, Month, "01", sep="-"), format="%Y-%m-%d"), y = cov_surg, color = factor(high_tn_rate))) +
  geom_line() +
  scale_color_discrete(name = "High TN rate") +
  labs(x = "Date", y = "Number of surgeries", title = "Surgeries by date and high TN rate")
```

## log transformation
```{r}
ccilog <- log(nsqip$cci + 1)
```

## Linear regression with log transformation
```{r}
mod1 <- lm(ccilog ~ age + sex + race + ... , data = nsqip)
```


## check the assumption of linear regression
```{r}
# Collinearity Diagnostic
vif(mod1)

# 


```

## if it fail, we will do medican regression

## EMM(covid variant) check

# Subgroup analysis
## Create the data for Delta and Omicron variant
## log transformation or median regression

# Sensitivity analysis
## regression with PS 
## IPW
## Standardization
























