---
title: "Homework0"
author: "Takuto Yoshida"
date: "2023-01-24"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Read in the nhefs.xlsx file from the EPI 289 course website. Show your log to demonstrate that the file was successfully assigned.  
```{r}
# Question1
install.packages("readxl", repos = "http://cran.us.r-project.org") # read excel file
library("readxl")
df <- read_excel("/Users/yoshidatakuto/Dropbox/HSPH/MPH-CLE/Spring/EPI289/Homework0/nhefs.xlsx")
```

2. Sort the data set by the variable seqn. Print out the ID number, age, and sex for the first 10 observations. 
```{r}
# Question2
## Sorting the data
df[1:10, c("seqn", "age", "sex")]
```
3. Find the mean systolic blood pressure and standard error for men and for women.
```{r}
# Question3
summary(df$sbp) # 77 missing values

# Separate the dataset by gender
df_men <- subset(df, sex==0)
df_women <- subset(df, sex==1)
summary(df_men$sbp) # 45 missing values
summary(df_women$sbp) # 32 missing values

# Find the mean and standard error of systolic blood pressure in men
mean_sbp_men <- mean(df_men$sbp, na.rm=T)
se_sbp_men <- sd(df_men$sbp, na.rm=T)/sqrt(length(df_men$sbp)-45)

# Find the mean and standard error of systolic blood pressure in women
mean_sbp_women <- mean(df_women$sbp, na.rm=T)
se_sbp_women <- sd(df_women$sbp, na.rm=T)/sqrt(length(df_women$sbp)-32)

# Print the results
cbind(Mean = mean_sbp_men, SE = se_sbp_men)
cbind(Mean = mean_sbp_women, SE = se_sbp_women)

```
4. What is the mean, 25th percentile, 50th percentile, 75th percentile, and interquartile range of weight in 1971 (in kilograms).
```{r}
# Question 4
summary(df$wt71) # No missing value
mean(df$wt71)
quantile(df$wt71, probs = c(0.25, 0.50, 0.75))
IQR(df$wt71)
```
5a. Using ifelse statements, create a new categorical variable corresponding to quartiles of weight in 1971 as based on the cut-points from Question (4). Give a tabulation of your results. 
```{r}
# Question 5a
## Create the new variables
df$wt71_cat <- ifelse(df$wt71<=59.65, 1, ifelse(df$wt71<=69.40, 2, ifelse(df$wt71<=79.95, 3, 4)))

## Tabulate the categorical variable
table(df$wt71_cat)                      
```

5b. Create quartiles for weight in 1971 using cut in R. Give a tabulation of your results. Do your results match those of Question (5a)? Why or why not?
```{r}
# Question 5b
df_quart <- cut(df$wt71, breaks = c(-Inf, quantile(df$wt71, probs = c(0.25)), quantile(df$wt71, probs = c(0.5)), quantile(df$wt71, probs = c(0.75)), Inf))

table(df_quart)
```

6. 	Using lm in R, fit a univariate linear regression model for the outcome weight in 1971 with number of cigarettes smoked per day in 1971 as the predictor. Report the parameter estimate for cigarettes smoked per day. 
```{r}
# Question 6
linear_model <- lm(df$wt71 ~ df$smokeintensity, data = df)
summary(linear_model)
coef(linear_model)
confint(linear_model)
```

7.   Create a cross-tabulation between sex and race.
```{r}
# Question 7
table(df$sex, df$race, dnn = c("sex", "race"))
```

8. 	Using lm in R, fit a multivariate linear regression model for the outcome weight in  1971 with age, sex, and race as the predictors. From this model, print the observed and predicted values of weight in 1971 for the first 5 observations. What is the predicted value of weight in 1971 for an individual of age 40, female, and of Black or other race/ethnicity? 
```{r}
# Question8
## Fit the multivariable linear regression model
df$sex <- as.factor(df$sex)
df$race <- as.factor(df$race)
adj_linear_model <- lm(wt71 ~ age + sex + race, data = df)

## Get the predicted values for the first 5 observations
predicted_values <- predict(adj_linear_model, newdata = df[1:5,])

## Get the observed value for the first 5 observations
observed_values <- df$wt71[1:5]

## Print the observed and predicted values
cbind(Observed = observed_values, Predicted = predicted_values)

# Predicted value
covariate_values <- data.frame(age = 40, sex = "1", race = "1")
predicted_values_q8 <- predict(adj_linear_model, newdata = covariate_values)
print(predicted_values_q8)
```

9. 	Fit the same model from Question (8) using glm in R and compare your results. 
```{r}
# Question9
adj_linear_model_2 <- glm(wt71 ~ age + sex + race, data = df)

## Get the predicted values for the first 5 observations
predicted_values_2 <- predict(adj_linear_model_2, newdata = df[1:5,])

## Get the observed value for the first 5 observations
observed_values <- df$wt71[1:5]

## Print the observed and predicted values
cbind(Observed = observed_values, Predicted = predicted_values_2)
```

10. Using glm with family specified as binomial in R, fit a multivariate logistic regression model for the outcome asthma diagnosis in 1971 with age, sex, race, and usual physical activity status (var active) as the predictors. Print the predicted probabilities of asthma diagnosis for the individuals with the first 5 ID numbers. 
```{r}
# Question10
## Fit the multivariable logistic regression model
df$asthma <- as.factor(df$asthma)
df$active <- as.factor(df$active)
adj_linear_model_3 <- glm(asthma ~ age + sex + race + active, data = df, family = binomial(link = 'logit'))

## Get the predicted values for the first 5 observations
predicted_values_3 <- predict(adj_linear_model_3, newdata = df[1:5,], type = "response")
predicted_values_3
```

11. (Optional) Create a graph that plots systolic blood pressure on the Y-axis and usual physical activity status (var active) on the X-axis.
```{r}
# Question 11
boxplot(sbp ~ active, data = df, xlab="Usual physical activity status", ylab="Systolic blood pressure", main="Q11 Answer")
```