---
title: "Homework1"
author: "Takuto Yoshida"
date: "2023-01-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part1

## Question13

```{r cars}
# read dataset
df <- read.csv("/Users/yoshidatakuto/Dropbox/HSPH/MPH-CLE/Spring/EPI289/Homework1/hmwk1 (1).csv")
str(df)
model1 <- glm(death ~ smk, family=binomial(link = 'logit'), data = df)
summary(model1)
coef(model1)
sd(df$smk)/sqrt(length(df$smk))
```

# Part2

## Question 30

```{r}
# Creating the interaction term
df$smdr <- df$smk * df$drink

# Fitting the model
model2 <- glm(death ~ smk + drink + smdr, family=binomial(link = 'logit'), data = df)
summary(model2)
coef(model2)
sd(df$smk)/sqrt(length(df$smk))
```

# Part3

## Question 48

```{r}
# Fitting the model
model3 <- glm(death ~ smk + overwt, family=binomial(link = 'logit'), data = df)
summary(model3)
coef(model3)
sd(df$smk)/sqrt(length(df$smk))
```

# Part4

## Question 52

```{r}
# Instrall Package
install.packages("readxl", repos = "http://cran.us.r-project.org") # read excel file
library("readxl")
df_nhefs <-read_excel("/Users/yoshidatakuto/Dropbox/HSPH/MPH-CLE/Spring/EPI289/Homework1/nhefs.xlsx")
str(df_nhefs)
df_nhefs$qsmk <- as.factor(df_nhefs$qsmk)
df_nhefs$sex <- as.factor(df_nhefs$sex)
df_nhefs$race <- as.factor(df_nhefs$race) # binary variable
df_nhefs$age_quad <- df_nhefs$age*df_nhefs$age
model4 <- glm(death ~ qsmk + sex + race + age + age_quad, data = df_nhefs, family=binomial('logit'))
summary(model4)
confint(model4, param="qsmk", level=.95)
exp(coefficients(model4))
exp(confint(model4))
```

# Question 57

```{r}
df_nhefs$smokeyrs_quad <- df_nhefs$smokeyrs*df_nhefs$smokeyrs
model5 <- glm(death ~ qsmk + sex + race + age + age_quad + smokeyrs + smokeyrs_quad, data = df_nhefs, family=binomial('logit'))
summary(model5)
confint(model5, param="qsmk", level=.95)
```

# Part5

## Question61

Generate a temporary dataset that excludes observations with missing values for weight gain and create the following categories of age (25-40, 41-60, and \>60).

```{r}
# Creating a temporary dataset excluding 
summary(df_nhefs$wt82_71) # 63 missing values
temp_df <- df_nhefs[!is.na(df_nhefs$wt82_71) & !is.na(df_nhefs$wtloss) & df_nhefs$wtloss==0,]

# Creating a new age category
summary(df_nhefs$age) # no missing value
temp_df$age_cat <- ifelse(temp_df$age >= 25 & temp_df$age <= 40, 0, ifelse(temp_df$age <= 60, 1, 2))
table(temp_df$age_cat)

```

## Question62
Use cut in R to generate categorical variables with 5, 10, 20 and 49 categories of age.
```{r}
# Check the summary of the data
summary(temp_df$age) # Min 25, Max 74

# Creating age category in 5 categories
temp_df$age_cat_5 <- cut(temp_df$age, breaks = seq(25, 74, by = (74-25)/5), labels = c(1:5))
table(temp_df$age_cat_5)

# Creating age category in 10 categories
temp_df$age_cat_10 <- cut(temp_df$age, breaks = seq(25, 74, by = (74-25)/10), labels = c(1:10))
table(temp_df$age_cat_10)

# Creating age category in 20 categories
temp_df$age_cat_20 <- cut(temp_df$age, breaks = seq(25, 74, by = (74-25)/20), labels = c(1:20))
table(temp_df$age_cat_20)

# Creating age category in 49 categories
temp_df$age_cat_49 <- cut(temp_df$age, breaks = seq(25, 74, by = (74-25)/49), labels = c(1:49))
table(temp_df$age_cat_49)
```

## Question63
Use plot in R to graphically explore the relation between age and weight gain.

```{r}
plot(temp_df$age, temp_df$wt82_71, xlab="Age", ylab="weight gain", main="Question 63 Answer")
```

## Question64
Fit a linear regression model of the form wt82_71= α0 + α1Age with age as a continuous variable and plot the values predicted by the model against the observed values.
```{r}
# Fitting the linear model
lm_mod <- lm(wt82_71 ~ age, data = temp_df)

# Plotting the values predicted by the model against the observed values
plot(temp_df$age, temp_df$wt82_71, xlab="Age", ylab="weight gain", main="Question 64 Answer")
abline(lm_mod, col = "red")
```

## Question 65.
Similarly, fit a linear regression model of the form wt82_71= α0 + α1Age + α2Age2 and plot the values predicted by the model against the observed values.
```{r}
# Fitting the linear model
lm_mod_2 <- lm(wt82_71 ~ age + I(age^2), data=temp_df)

# plotting the values predicted by the model against the observed values.
plot(temp_df$age, temp_df$wt82_71, xlab="Age", ylab="weight gain", main="Question 65 Answer")
abline(lm_mod_2, col="red")
```

# Question 66.
Fit a linear regression model including age in 5, 10, 20 and 49 categories and plot the values predicted by the model against the observed values.
```{r}
# Fitting the moderl by using age in 5 category and plot the predicted and observed values
lm_mod_3 <- lm(wt82_71 ~ age_cat_5, data=temp_df)
predicted_value_5 <- predict(lm_mod_3, temp_df)
plot(temp_df$wt82_71, predicted_value_5)

# Fitting the moderl by using age in 10 category and plot the predicted and observed values
lm_mod_4 <- lm(wt82_71 ~ age_cat_10, data=temp_df)
predicted_value_10 <- predict(lm_mod_4, temp_df)
plot(temp_df$wt82_71, predicted_value_10)

# Fitting the moderl by using age in 20 category and plot the predicted and observed values
lm_mod_5 <- lm(wt82_71 ~ age_cat_20, data=temp_df)
predicted_value_20 <- predict(lm_mod_5, temp_df)
plot(temp_df$wt82_71, predicted_value_20)

# Fitting the moderl by using age in 49 category and plot the predicted and observed values
lm_mod_6 <- lm(wt82_71 ~ age_cat_49, data=temp_df)
predicted_value_49 <- predict(lm_mod_6, temp_df)
plot(temp_df$wt82_71, predicted_value_49)
```















